---
title: "Untitled"
author: "Harly Durbin"
date: "8/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(readr)
library(dplyr)
library(stringr)

source(here::here("source_functions/animal_table.R"))
source(here::here("source_functions/iterative_id_search.R"))
```

# Setup 

```{r}
animal_table <- get_animal_table("200806")
```


```{r}
xref <- 
  read_csv("~/Downloads/xref.csv") %>% 
  mutate(sim_breed = str_extract(animal, "^[[:upper:]]{3}"),
         sim_sex = str_extract(animal, "(?<=^[[:upper:]]{6})M|F"),
         sim_iid = glue::glue("SIMUSA{sim_sex}{str_pad(asa_nbr, 12, side = 'left', pad = '0')}")) 
```

# Find matches

```{r}

matching <-
  xref %>%
  id_search(
    source_col = animal,
    search_df = animal_table,
    search_col = international_id,
    key_col = Lab_ID,
    match_source = TRUE
  ) %>%
  id_search(
    source_col = sim_iid,
    search_df = animal_table,
    search_col = international_id,
    key_col = Lab_ID,
    match_source = TRUE
  ) %>%
  id_search(
    source_col = asa_nbr,
    search_df = animal_table %>%
      filter(BC == "SIM") %>% 
      rename(Reg_SIM = Reg),
    search_col = Reg_SIM,
    key_col = Lab_ID,
    match_source = TRUE
  ) %>%
  id_search(
    source_col = asa_nbr,
    search_df = animal_table %>%
      filter(BC == "SIM") %>% 
      mutate(Reg = str_pad(Reg, width = 7, side = "left", pad = "0")) %>% 
      rename(Reg_SIM_pad = Reg),
    search_col = Reg_SIM_pad,
    key_col = Lab_ID,
    match_source = TRUE
  ) %>% 
  left_join(
    animal_table %>%
      select(
        Lab_ID,
        mu_reg = Reg,
        mu_breed = BC,
        mu_sex = Sex,
        duplicate,
        mu_comment = Comment,
        mu_iid = international_id
      )
  ) %>% 
  mutate(mu_sex = if_else(mu_sex == "U", sim_sex, mu_sex))
```

# Tally by `match_source`

```{r}
matching %>% 
  distinct(asa_nbr, match_source) %>% 
  group_by(match_source) %>% 
  tally(sort = TRUE)
```

```{r}
matching %>% 
  filter(match_source == "asa_nbr-Reg_SIM_pad") %>% 
  select(animal, mu_ia, everything())
```

# Duplicates

```{r}
duplicates <-
  matching %>% 
  group_by(asa_nbr) %>% 
  filter(n() > 1) %>% 
  ungroup() %>% 
  arrange(asa_nbr) %>% 
  select(asa_nbr, animal, match_source, duplicate, Lab_ID)
```

# Sex discrepancies 

```{r}
wrong_sex <-
  matching %>% 
  filter(sim_sex != mu_sex) %>% 
  select(asa_nbr, animal, match_source, Sex = mu_sex, Lab_ID)
```


```{r}
matching %>% 
  filter(sim_sex == mu_sex) %>% 
  filter(str_detect(match_source, "asa_nbr-Reg"))
```

# Unmatched

```{r}
unmatched <-
  matching %>% 
  filter(is.na(Lab_ID)) %>% 
  select(asa_nbr, animal, match_source, Lab_ID)
  
```

# Confident

```{r}
confident <-
  matching %>% 
  filter(!is.na(Lab_ID)) %>% 
  filter(sim_sex == mu_sex) %>% 
  group_by(asa_nbr) %>% 
  filter(n() == 1) %>% 
  ungroup() %>% 
  select(asa_nbr, animal, match_source, Lab_ID)
```

```{r}
confident %>% 
  group_by(match_source) %>% 
  tally(sort = TRUE)
```


# How many `sim_ia-international_id` matches associated with hair shedding?

```{r}
matching %>% 
  filter(match_source == "sim_iid-international_id") %>% 
  group_by(mu_comment) %>% 
  tally(sort = TRUE)
```


# Export

I found lab IDs for about half of these I think (see attached). I added a column for the `match_source` (which column in the "[ASA data column]-[column in our system]) in case that's helpful in deciding what to believe and what not to believe. 

Tabs in the file:
* `confident`
    * 95,383 matches based on a direct match of the international ID provided by ASA and the international ID in our system
    * 4,187 matches based on a generated international ID ("SIMUSA" + sex inferred from the provided `animal` column + registration number provided in `asa_nbr` column padded with zeroes to 12 digits)
    * 539 matches based on the registration number provided in the `asa_nbr` column and the `Reg` column in our system, subset to rows where `BC` = `SIM` OR subset to rows where `BC` = `SIM` and padded to 7 digits with zeroes like in the ASA file. Eyeballing these, it looks like most or all are probably true matches but I think that's your call
* `duplicates`: 723 animals that match up to 2 or more Lab_IDs (1,470 total duplicates). It appears that most of these are known duplicates.
* `wrong_sex`: 299 animals that match up to a lab ID, but don't match the sex we have recorded in our system (excluding those where the sex we have is "U")
* `unmatched`: 115,814 animals without a lab ID based on the above criteria

```{r}
writexl::write_xlsx(list("confident" = confident, "duplicates" = duplicates, "wrong_sex" = wrong_sex, "unmatched" = unmatched), path = "~/Desktop/200806.ASA_Lab_ID.xlsx")
```

