---
title: "Simmental Animal Locations"
author: "Troy Rowan"
date: "2020-07-31"
output: 
  html_document:
    code_folding: show
editor_options:
  chunk_output_type: inline
---

```{r, include = FALSE}
set.seed(325333)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(workflowr)
library(knitr)
library(rgdal)
library(raster)
library(fpc)
library(RStoolbox)
library(maps)
library(ggthemes)
library(cowplot)
library(purrr)
library(factoextra)
source(here::here("code", "map_functions.R"))
library(corrr)
library(ggcorrplot)
library(factoextra)
library(here)
library(tidyverse)
setwd("C:/Users/troyrowan/Documents/GitHub/local_adaptation_sequence")
```


### Reading in phenotypes, pedigree, etc. 

```{r}

sim_animals = 
  read_csv("data/mizzou-data-request/perf.csv") %>% 
  left_join(read_csv("data/mizzou-data-request/xref.csv")) %>% 
  rename(international_id = animal) 


sim_pedigree = 
  read_csv("data/mizzou-data-request/6-gen-ped.csv")
sim_xref = 
  read_csv("data/mizzou-data-request/xref.csv")

zip_info = 
  read_csv("data/uszips.csv") %>% 
  dplyr::select(zip, x = lng, y = lat, city, state_id)%>% 
  mutate(lat = round(y, 1),
         long = round(x, 1)) %>% 
  select(-x, -y)

k9_all_s1 = readRDS("output/k9.allvars.seed1.rds") #Breaks out Fescue Belt
k9_all_s2 = readRDS("output/k9.allvars.seed2.rds") #Similar fescue belt to 3-var 
k9_three_s1 = readRDS("output/k9.threevars.seed1.rds") #These are basically equivalent, just with different seeds
k9_three_s2 = readRDS("output/k9.threevars.seed2.rds")
k10 = readRDS("output/k10.allvars.seed2.rds") #Ten variable 
```
### Pulling lab IDs for genotype dump
This is from the animal table as of August 21st, 2020
Harly turned this into an RDS file using one of her scripts and weird suite of packages (almost broke her computer)
This behaved a bit curiously as I kept dropping ~10 K animals from the Simmental data sheet that I read in. Turns out that they had repeated records for mature weight, so the same cow could be listed multiple times. Upon filtering those out and matching as many animals as possible to a Reg or Ref_ID based on their ASA Registration number, I get 100,559 distinct lab IDs. They're written out and sent to Bob for a data dump on CIFS prior to imputation on Lewis. 

There are 99,932 individuals that are adequately accoutned for in the database

```{r}
animal_table = 
  readRDS("data/animal_table.rds")
SIM =
  rbind(
    animal_table %>% 
      filter(Reg %in% sim_animals$asa_nbr), 
    animal_table %>% 
      filter(!is.na(Ref_ID) & Ref_ID %in% sim_animals$asa_nbr), 
    animal_table %>% 
      filter(!is.na(Ref_ID2) & Ref_ID2 %in% sim_animals$asa_nbr),
    animal_table %>% 
      filter(!is.na(Ref_ID3) & Ref_ID3 %in% sim_animals$asa_nbr))

#This showed me animals that have multiple entries in the "sim_animal" file 
#sim_animals %>% group_by(asa_nbr) %>% count(sort = TRUE)
#filter(sim_animals, !asa_nbr %in% c(SIM$Reg, SIM$Ref_ID, SIM$Ref_ID2, SIM$Ref_ID3))

SIM %>% 
  select(Lab_ID) %>% 
  distinct() %>%
  write_csv("output/200822_Lab_IDs.csv")



```



### Completeness of data

98,335 individuals (unique ASA reg numbers) match up to entries in the database

Important to note that many of these are repeats caused by mature weight phenotypes being different for a single animal (up to 7 repeated records).

This does, however show the number of BW, WW, and YW that we have access to 
```{r}


sim_animals %>% 
  select(BW = bw, WW = ww, YW = yw, MW = mw) %>% 
 summarise_all(funs(sum(!is.na(.)))) %>% 
  kable()


sim_animals %>% 
  filter(asa_nbr %in% c(SIM$Reg, SIM$Ref_ID, SIM$Ref_ID2, SIM$Ref_ID3)) %>% 
  select(asa_nbr) %>%
  unique() %>% 
  count() %>% .$n %>% 
  paste(., "individuals in MU database")
```


## Simmental Location Counts {.tabset}

### K=9 All Variables (Seed #1)

```{r}
k9_zips = 
  left_join(zip_info, k9_all_s1 %>% 
    mutate(lat = round(y,1),
           long = round(x, 1)),
    by = c("lat", "long")) %>% 
    select(-x, -y) %>% 
    filter(!is.na(layer)) %>% 
    distinct() %>% 
  mutate(region = 
           case_when(
            layer == 5 ~ "FescueBelt",
            layer == 3 ~ "Southeast",
            layer == 6 ~ "ForestedMountains",
            layer == 9 ~ "Desert",
            layer == 8 ~ "AridPrairie",
            layer == 4 ~ "CornBelt",
            layer == 7 ~ "UpperMidwest",
            layer == 2 ~ "Foothills",
            layer == 1 ~ "HighPlains"
          )
         ) %>% 
  rename(zone = layer)



sim_animals %>% 
  #mutate(zip = as.numeric(breeder_zip)) %>% 
  filter(breeder_zip %in% k9_zips$zip) %>% 
  select(asa_nbr, breeder_zip, international_id) %>% 
  left_join(k9_zips, by = c("breeder_zip" = "zip")) %>% 
  group_by(region) %>% 
  count() %>% kable()

```

### K=9 All Variables (Seed #2)
Here, Rainforest becomes a region, but isn't counted here when we do K=9 with all variables and this seed
```{r}
k9_zip_s2 = 
  left_join(zip_info, k9_all_s2 %>% 
    mutate(lat = round(y,1),
           long = round(x, 1)),
    by = c("lat", "long")) %>% 
    select(-x, -y) %>% 
    filter(!is.na(layer)) %>% 
    distinct() %>% 
  mutate(region = 
           case_when(
            layer == 7 ~ "Fescue Belt",
            layer == 9 ~ "Southeast",
            layer == 2 ~ "Forested Mountains",
            layer == 6 ~ "Desert",
            layer == 8 ~ "Arid Prairie",
            layer == 4 ~ "Rainforest",
            layer == 3 ~ "Upper Midwest",
            layer == 5 ~ "Foothills",
            layer == 1 ~ "High Plains"
          )
         ) %>% 
  rename(zone = layer)


sim_animals %>% 
  #mutate(zip = as.numeric(breeder_zip)) %>% 
  filter(breeder_zip %in% k9_zip_s2$zip) %>% 
  select(asa_nbr, breeder_zip, international_id) %>%  
  left_join(k9_zip_s2, by = c("breeder_zip" = "zip")) %>% 
  group_by(region) %>% 
  count() %>% kable()
```

### K=9 Three Variables
Seed doesn't appear to cause any issues with major region changes so far as I can see.

```{r}
k9_zips_threevar = 
  left_join(zip_info, k9_three_s1 %>% 
    mutate(lat = round(y,1),
           long = round(x, 1)),
    by = c("lat", "long")) %>% 
    select(-x, -y) %>% 
    filter(!is.na(layer)) %>% 
    distinct() %>% 
  mutate(region = 
           case_when(
            layer == 1 ~ "Fescue Belt",
            layer == 2 ~ "Southeast",
            layer == 9 ~ "Forested Mountains",
            layer == 4 ~ "Desert",
            layer == 5 ~ "Arid Prairie",
            layer == 8 ~ "Rainforest",
            layer == 6 ~ "Upper Midwest",
            layer == 3 ~ "Foothills",
            layer == 7 ~ "High Plains"
          )
         ) %>% 
  rename(zone = layer)



sim_animals %>% 
  #mutate(zip = as.numeric(breeder_zip)) %>% 
  filter(breeder_zip %in% k9_zips_threevar$zip) %>% 
  select(asa_nbr, breeder_zip, international_id) %>%  
  left_join(k9_zips_threevar, by = c("breeder_zip" = "zip")) %>% 
  group_by(region) %>% 
  count() %>% kable()
```



## Simmental Location Maps

### K = 9 Map

```{r, fig.height=4, fig.width=8}
sim_animals %>% 
  #mutate(zip = as.numeric(breeder_zip)) %>% 
  filter(breeder_zip %in% k9_zips$zip) %>% 
  left_join(k9_zips, by = c("breeder_zip" = "zip")) %>%
  #rename(zone = layer) %>% 
  mapplotting(max = 20)
```



