---
title: "Selection Manuscript Figures"
author: "Troy Rowan"
date: "2020-09-20"
output: 
  html_document:
    code_folding: show
editor_options:
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

```{r, setup, include = FALSE}
set.seed(325333)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = here::here())
library(workflowr)
library(knitr)
library(maps)
library(ggthemes)
library(purrr)
library(factoextra)
library(corrr)
library(ggcorrplot)
library(factoextra)
library(here)
library(tidyverse)
library(lubridate)
library(pedigree)
library(qvalue)
library(GALLO)
library(cowplot)
library(gprofiler2)
library(DT)
library(ggforce)
library(vroom)
#source("../code/annotation_functions.R")
```

```{r}
source("code/GCTA_functions.R")
source("code/annotation_functions.R")
simmental = read_csv("output/200907_SIM/phenotypes/200907_SIM.info.csv") %>% 
  left_join(read_tsv("output/200907_SIM/phenotypes/200907_SIM.generation_proxy.txt", col_names = FALSE) %>% 
    select(international_id = X1, birth_year = X4))
redangus = read_csv("output/200910_RAN/phenotypes/200910_RAN.info.csv")
```

# Figure 1 

Reading in files
```{r}
ran_seq_gpsm_age = 
  seq(1,29) %>% 
    purrr::map(~read_gwas(paste0("output/200910_RAN/seq_gwas/200910_RAN.age.chr",.x,".mlma.gz"))%>% 
                 filter(p<0.01)) %>% 
    purrr::reduce(bind_rows)

#read_gwas("output/200910_RAN/gwas/200910_RAN.age.850K.mlma.gz")

ran_gpsm_sigsnps =
  ran_seq_gpsm_age %>% 
  filter(p < 4.289e-9) %>% 
  .$SNP

ran_seq_gpsm_young_age = 
  seq(1,29) %>% 
    purrr::map(~read_gwas(paste0("output/200910_RAN/seq_gwas/200910_RAN.young_age.chr",.x,".mlma.gz")) %>% 
                 filter(p < 0.01)) %>% 
    purrr::reduce(bind_rows) %>% 
  filter(p != 0)

sim_seq_gpsm_purebred_age = 
  seq(1,29) %>% 
    purrr::map(~read_gwas(paste0("output/200907_SIM/seq_gwas/200907_SIM.pb_age.chr",.x,".mlma.gz")) %>% 
                 filter(p<0.01)) %>% 
    purrr::reduce(bind_rows)

sim_seq_gpsm_age = 
  seq(1,29) %>% 
    purrr::map(~read_gwas(paste0("output/200907_SIM/seq_gwas/200907_SIM.full_age.chr",.x,".mlma.gz")) %>% 
                 filter(p<0.01)) %>% 
    purrr::reduce(bind_rows)

ran_age_seq_cojo = 
  list.files(path='output/200910_RAN/seq_cojo/',pattern="200910_RAN.age.chr") %>% 
  purrr::map(
  ~read_tsv(paste0("output/200910_RAN/seq_cojo/",.x),
           col_names = c("Chr", "SNP", "bp", "freq", "refA", "b", "se", "p", "n", "freq_geno", "bJ", "bJ_se", "pJ", "LD_r"),
           skip = 1) %>% 
    mutate(chrbp = paste(Chr, bp, sep = ":")) %>% 
    select(SNP, CHR = Chr, BP = bp, p, pJ, chrbp)) %>% 
    purrr::reduce(bind_rows) %>% 
  filter(p < 5e-8)

ran_old_seq_cojo = 
  list.files(path='output/200910_RAN/seq_cojo/',pattern="200910_RAN.old_age") %>% 
  purrr::map(
  ~read_tsv(paste0("output/200910_RAN/seq_cojo/",.x),
           col_names = c("Chr", "SNP", "bp", "freq", "refA", "b", "se", "p", "n", "freq_geno", "bJ", "bJ_se", "pJ", "LD_r"),
           skip = 1) %>% 
    select(SNP, CHR = Chr, BP = bp, p, pJ)) %>% 
    purrr::reduce(bind_rows) %>% 
  filter(p < 5e-8)

ran_young_seq_cojo = 
  list.files(path='output/200910_RAN/seq_cojo/',pattern="200910_RAN.young_age.chr") %>% 
  purrr::map(
  ~read_tsv(paste0("output/200910_RAN/seq_cojo/",.x),
           col_names = c("Chr", "SNP", "bp", "freq", "refA", "b", "se", "p", "n", "freq_geno", "bJ", "bJ_se", "pJ", "LD_r"),
           skip = 1) %>% 
    mutate(chrbp = paste(Chr, bp, sep = ":")) %>% 
    select(SNP, CHR = Chr, BP = bp, p, pJ, chrbp)) %>% 
    purrr::reduce(bind_rows) %>% 
  filter(p < 5e-8)

sim_pb_seq_cojo = 
  list.files(path='output/200907_SIM/seq_cojo/',pattern="200907_SIM.pb_age.chr") %>% 
  purrr::map(
  ~read_tsv(paste0("output/200907_SIM/seq_cojo/",.x),
           col_names = c("Chr", "SNP", "bp", "freq", "refA", "b", "se", "p", "n", "freq_geno", "bJ", "bJ_se", "pJ", "LD_r"),
           skip = 1)  %>% 
    mutate(chrbp = paste(Chr, bp, sep = ":")) %>% 
    select(SNP, CHR = Chr, BP = bp, p, pJ, chrbp)) %>% 
    purrr::reduce(bind_rows)%>% 
  filter(p < 5e-8)

sim_seq_cojo = 
  list.files(path='output/200907_SIM/seq_cojo/',pattern="200907_SIM.full_age") %>% 
  purrr::map(
  ~read_tsv(paste0("output/200907_SIM/seq_cojo/",.x),
           col_names = c("Chr", "SNP", "bp", "freq", "refA", "b", "se", "p", "n", "freq_geno", "bJ", "bJ_se", "pJ", "LD_r"),
           skip = 1)  %>% 
    mutate(chrbp = paste(Chr, bp, sep = ":")) %>% 
    select(SNP, CHR = Chr, BP = bp, p, pJ, chrbp)) %>% 
    purrr::reduce(bind_rows) %>% 
  filter(p < 5e-8)


```




Red Angus GPSM

```{r}
plot_grid(
  ran_seq_gpsm_age %>% 
    ggmanhattan2(sig_threshold_p = 4.289e-9,
                 sigsnps = ran_age_seq_cojo$SNP,
                 sigsnpcolor = "dodgerblue")+
    geom_hline(yintercept = 7.3, color = "blue", size = 0.25),
  ran_seq_gpsm_age %>% 
    ggmanhattan2(sig_threshold_p = 4.289e-9,
                 sigsnps = ran_age_seq_cojo$SNP,
                 sigsnpcolor = "dodgerblue")+
    geom_hline(yintercept = 7.3, color = "blue", size = 0.25)+
    ylim(c(2,15)),
  ran_seq_gpsm_young_age %>% 
      ggmanhattan2(sig_threshold_p = 4.289e-9,
                 sigsnps = ran_young_seq_cojo$SNP,
                 sigsnpcolor = "indianred")+
      geom_hline(yintercept = 7.3, color = "blue", size = 0.25),
  ran_seq_gpsm_young_age %>% 
    ggmanhattan2(sig_threshold_p = 4.289e-9,
                 sigsnps = ran_young_seq_cojo$SNP,
                 sigsnpcolor = "indianred")+
    geom_hline(yintercept = 7.3, color = "blue", size = 0.25)+
    ylim(c(2,15)),
    nrow = 2, labels = c("A", "B", "C", "D"))

ggsave("output/figures/RAN_GPSM.png",
       width = 12, height = 5, units = "in")

ran_age_sigsnps = filter(ran_seq_gpsm_age, p < 5e-8) %>% .$SNP
ran_young_age_sigsnps = filter(ran_seq_gpsm_young_age, p < 5e-8) %>% .$SNP
plot_grid(
  ran_seq_gpsm_age %>% 
    ggmanhattan2(sig_threshold_p = 4.289e-9,
                 sigsnps = ran_young_age_sigsnps,
                 sigsnpcolor = "dodgerblue")+
    geom_hline(yintercept = 7.3, color = "blue"),
  ran_seq_gpsm_age %>% 
    ggmanhattan2(sig_threshold_p = 4.289e-9,
                 sigsnps = ran_young_age_sigsnps,
                 sigsnpcolor = "dodgerblue")+
    geom_hline(yintercept = 7.3, color = "blue")+
    ylim(c(2,15)),
  ran_seq_gpsm_young_age %>% 
      ggmanhattan2(sig_threshold_p = 4.289e-9,
                 sigsnps = ran_age_sigsnps,
                 sigsnpcolor = "indianred")+
      geom_hline(yintercept = 7.3, color = "blue"),
  ran_seq_gpsm_young_age %>% 
    ggmanhattan2(sig_threshold_p = 4.289e-9,
                 sigsnps = ran_age_sigsnps,
                 sigsnpcolor = "indianred")+
    geom_hline(yintercept = 7.3, color = "blue")+
    ylim(c(2,15)),
    nrow = 2, labels = c("A", "B", "C", "D"))
```

Simmental GPSM

```{r}

plot_grid(
  sim_seq_gpsm_purebred_age %>% 
    ggmanhattan2(sig_threshold_p = 4.289e-9,
                 sigsnps = sim_pb_seq_cojo$SNP,
                 sigsnpcolor = "slateblue3")+
    geom_hline(yintercept = 7.3, color = "blue", size = 0.25),
  sim_seq_gpsm_purebred_age %>% 
    ggmanhattan2(sig_threshold_p = 4.289e-9,
                 sigsnps = sim_pb_seq_cojo$SNP,
                 sigsnpcolor = "slateblue3")+
    geom_hline(yintercept = 7.3, color = "blue", size = 0.25)+
    ylim(c(2,15)),
  sim_seq_gpsm_age %>% 
      ggmanhattan2(sig_threshold_p = 4.289e-9,
                 sigsnps = sim_seq_cojo$SNP,
                 sigsnpcolor = "springgreen3")+
      geom_hline(yintercept = 7.3, color = "blue", size = 0.25),
  sim_seq_gpsm_age %>% 
    ggmanhattan2(sig_threshold_p = 4.289e-9,
                 sigsnps = sim_seq_cojo$SNP,
                 sigsnpcolor = "springgreen3")+
    geom_hline(yintercept = 7.3, color = "blue", size = 0.25)+
    ylim(c(2,15)),
    labels = c("A", "B", "C", "D"),
    nrow = 2)

ggsave("output/figures/SIM_GPSM.png",
       width = 12, height = 5, units = "in")

filter(sim_seq_gpsm_purebred_age, p < 4.289e-9) %>% .$SNP %>% length()

plot_grid(
  sim_seq_gpsm_purebred_age %>% 
    ggmanhattan2(sig_threshold_p = 4.289e-9,
                 sigsnps = filter(sim_seq_gpsm_age, p < 4.289e-9) %>% .$SNP,
                 sigsnpcolor = "slateblue3")+
    geom_hline(yintercept = 7.3, color = "blue"),
  sim_seq_gpsm_purebred_age %>% 
    ggmanhattan2(sig_threshold_p = 4.289e-9,
                 sigsnps = filter(sim_seq_gpsm_age, p < 4.289e-9) %>% .$SNP,
                 sigsnpcolor = "slateblue3")+
    geom_hline(yintercept = 7.3, color = "blue")+
    ylim(c(2,15)),
  sim_seq_gpsm_age %>% 
      ggmanhattan2(sig_threshold_p = 4.289e-9,
                 sigsnps = filter(sim_seq_gpsm_purebred_age, p < 4.289e-9) %>% .$SNP,
                 sigsnpcolor = "springgreen3")+
      geom_hline(yintercept = 7.3, color = "blue"),
  sim_seq_gpsm_age %>% 
    ggmanhattan2(sig_threshold_p = 4.289e-9,
                 sigsnps = filter(sim_seq_gpsm_purebred_age, p < 4.289e-9) %>% .$SNP,
                 sigsnpcolor = "springgreen3")+
    geom_hline(yintercept = 7.3, color = "blue")+
    ylim(c(2,15)),
    nrow = 2)
ggsave("output/figures/SIM_GPSM2.png",
       width = 12, height = 6, units = "in")
```

## Chromosome 23

AF trajectory plotting: Read in files make plot in this collapsed cell
These are from the sequence data, extracted COJO SNPs for FULL and YOUNG datasets
```{r}
by2 <- function(n){
  x = n/2
  return(x)
}


ran_full_af = 
  read_delim("output/200910_RAN/seq_cojo/200910_RAN.age.cojosnps.raw",
             delim = " ")%>%
  filter(!str_detect(FID, "NA")) %>%
  dplyr::rename(birthyear = FID) %>%
  mutate_at(.vars = vars(contains(":")), .funs = by2) %>%
  select(-PAT, -MAT, -SEX, -PHENOTYPE) %>%
  reshape2::melt(id = c("birthyear", "IID")) %>%
  mutate(rs = str_split_fixed(variable, "_", n = 2)[,1],
         AF = value) %>% 
  select(-value, -IID, -variable) %>% 
  filter(rs %in% ran_age_seq_cojo$SNP)

ran_young_af = 
  read_delim("output/200910_RAN/seq_cojo/200910_RAN.young_age.cojosnps.raw",
             delim = " ") %>%
  filter(!str_detect(FID, "NA")) %>%
  dplyr::rename(birthyear = FID) %>%
  mutate_at(.vars = vars(contains(":")), .funs = by2) %>%
  select(-PAT, -MAT, -SEX, -PHENOTYPE) %>%
  reshape2::melt(id = c("birthyear", "IID")) %>%
  mutate(rs = str_split_fixed(variable, "_", n = 2)[,1],
         AF = value) %>% 
  select(-value, -IID, -variable)%>% 
  filter(rs %in% ran_young_seq_cojo$SNP)


sim_af = 
  read_delim("output/200907_SIM/af_trajectories/200907_SIM.chr23.af.raw.gz",
             delim = " ") %>%
  left_join(select(simmental, international_id, birth_year), by = c("IID" = "international_id")) %>% 
  select(birthyear =birth_year, IID, everything()) %>%
  select(-FID) %>% 
  filter(!is.na(birthyear)) %>%
  mutate_at(.vars = vars(contains(":")), .funs = by2) %>%
  select(-PAT, -MAT, -SEX, -PHENOTYPE) %>%
  reshape2::melt(id = c("birthyear", "IID")) %>%
  mutate(rs = str_split_fixed(variable, "_", n = 2)[,1],
         AF = value) %>% 
  select(-value, -IID, -variable)

chr23_af = 
  bind_rows(
    ran_full_af %>% 
      filter(rs %in% ran_age_seq_cojo$SNP) %>% 
      mutate(CHR = as.numeric(str_split_fixed(rs, ":", n = 4)[,1]),
             BP = as.numeric(str_split_fixed(rs, ":", n = 4)[,2])) %>% 
      filter(CHR == 23) %>% 
      left_join(ran_age_seq_cojo, by = c("rs" = "SNP")) %>% 
      mutate(analysis = "RedAngus"),
    #   filter(rs %in% c("23:1032665:T:C", "23:1324051:C:T", "23:1366335:G:C", "23:1366691:G:T", "23:1768070:T:C", "23:1783688:T:C", "23:1805412:A:G")) %>% 
    #   mutate(analysis = "Full Red Angus") %>% 
    # left_join(ran_age_seq_cojo, by = c("rs" = "SNP")),
    # ran_young_af %>% 
    #   filter(rs %in% c("23:1324051:C:T","23:1366335:G:C","23:1366691:G:T", "23:1768070:T:C", "23:1800642:A:G", "23:1832160:A:G", "23:1918909:T:A")) %>% 
    #   mutate(analysis = "Red Angus")%>% 
    sim_af %>% 
      filter(rs %in% sim_seq_cojo$SNP) %>% 
      mutate(CHR = as.numeric(str_split_fixed(rs, ":", n = 4)[,1]),
             BP = as.numeric(str_split_fixed(rs, ":", n = 4)[,2])) %>% 
      filter(CHR == 23) %>%  
      mutate(analysis = "Simmental")%>% 
    left_join(sim_seq_cojo, by = c("rs" = "SNP"))
  )


```

```{r}
chr2_af %>%
  left_join(ran_age_seq_cojo, by = c("rs" = "SNP")) %>% 
  filter(str_detect(rs, "^2:")) %>% 
  mutate(p = case_when(p == 0 ~ 1e-310,
                        TRUE ~ p)) %>%
  # filter(analysis == "Full Red Angus" | rs %in% c("2:53151542:A:G", "2:103217345:T:A", "2:7500278:A:G", "2:22339424:A:G", "2:8100524:G:T", "2:8674849:T:C", "2:60054014:G:A")) %>%
  ggplot(aes(x = birthyear, y = AF, group = rs, color = -log10(p)))+
  geom_smooth(se = FALSE)+
  facet_grid(~analysis)+
  theme_cowplot()+
  ylim(c(0,0.5))+
  scale_color_viridis_c()+
  labs(x = "Birth Year", y = "Allele Frequency", color = "-log10(p)")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

chr23_af %>% 
  mutate(p = case_when(p == 0 ~ 1e-310,
                        TRUE ~ p)) %>% 
  ggplot(aes(x = birthyear, y = AF, group = rs, color = -log10(p)))+
  geom_smooth(se = FALSE, size = 2)+
  facet_grid(~analysis,)+
  theme_cowplot()+
  ylim(c(0,1))+
  scale_color_viridis_c(option = "inferno", begin = 0.25, end = 0.9)+
  labs(x = "Birth Year", y = "Allele Frequency", color = "-log10(p)")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```


```{r}
CHR_YGPSM = read_tsv("output/200910_RAN/seq_gwas/200910_RAN.young_age.chr23.mlma.gz")
CHR_YCOJO = ran_young_seq_cojo %>% filter(CHR == 23)

CHR_GPSM = read_tsv("output/200910_RAN/seq_gwas/200910_RAN.age.chr23.mlma.gz")
CHR_COJO = ran_age_seq_cojo %>% filter(CHR == 23)

CHR_SGPSM = read_tsv("output/200907_SIM/seq_gwas/200907_SIM.full_age.chr23.mlma.gz")
CHR_SCOJO = sim_seq_cojo %>% filter(CHR == 23)

CHR_RAISD = read_tsv("output/200910_RAN/sfs_selection/RAiSD_Report.200910_RAN.chr23.gz",
                 skip = 1,
                 col_names = c("BP", "stat", "finish", "VAR", "SFS", "LD", "Mu"))
CHR_NSL = read_tsv("output/selscan_window/200907_SIM.all.nsl.out.100bins.norm.windows.txt",
                   col_types = cols(SNP = col_character(), window = col_character())) %>% filter(CHR == 23)

CHR_YGPSM_CHIP = read_tsv("output/200910_RAN/gwas/200910_RAN.young_age.850K.mlma.gz") %>% 
  filter(Chr == 23)


ATAC = 
  FAANG %>% 
    filter(CHR == 23,
           START >= 1000000,
           START <= 1900000) %>% 
  mutate(start = START/1000000,
         stop = STOP/1000000,
         SNP = paste(CHR, (START+STOP)/2))

functional_plot = 
  ggplot()+
    geom_rect(data=ATAC %>% filter(TISSUE == "Adipose"), mapping=aes(x = NULL, y = NULL, xmin=start, xmax=stop, ymin=14, ymax=16, fill = MARK))+ 
    geom_text(aes(x = 2, y = 15, label = "Adipose"))+
    geom_rect(data=ATAC %>% filter(TISSUE == "Cerebellum"), mapping=aes(x = NULL, y = NULL, xmin=start, xmax=stop, ymin=12, ymax=14, fill = MARK))+
    geom_text(aes(x = 2, y = 13, label = "Cerebellum"))+
    geom_rect(data=ATAC %>% filter(TISSUE == "Cortex"), mapping=aes(x = NULL, y = NULL, xmin=start, xmax=stop, ymin=10, ymax=12, fill = MARK))+
    geom_text(aes(x = 2, y = 11, label = "Cortex"))+
    geom_rect(data=ATAC %>% filter(TISSUE == "Hypothalamus"), mapping=aes(x = NULL, y = NULL, xmin=start, xmax=stop, ymin=10, ymax=8, fill = MARK))+
    geom_text(aes(x = 2, y = 9, label = "Hypothalamus"))+
    geom_rect(data=ATAC %>% filter(TISSUE == "Liver"), mapping=aes(x = NULL, y = NULL, xmin=start, xmax=stop, ymin=6, ymax=8, fill = MARK))+
    geom_text(aes(x = 2, y = 7, label = "Liver"))+
    geom_rect(data=ATAC %>% filter(TISSUE == "Lung"), mapping=aes(x = NULL, y = NULL, xmin=start, xmax=stop, ymin=2, ymax=6, fill = MARK))+
    geom_text(aes(x = 2, y = 5, label = "Lung"))+
    geom_rect(data=ATAC %>% filter(TISSUE == "Muscle"), mapping=aes(x = NULL, y = NULL, xmin=start, xmax=stop, ymin=2, ymax=4, fill = MARK))+
    geom_text(aes(x = 2, y = 3, label = "Muscle"))+
    geom_rect(data=ATAC %>% filter(TISSUE == "Spleen"), mapping=aes(x = NULL, y = NULL, xmin=start, xmax=stop, ymin=0, ymax=2, fill = MARK))+
    geom_text(aes(x = 2, y = 1, label = "Spleen"))+
    #geom_rect(aes(x = NULL, y = NULL, xmin=0.268191, xmax=0.961226, ymin=-12, ymax=-2), fill = "indianred")+
    geom_rect(aes(x = NULL, y = NULL, xmin=1.448844, xmax=1.456438, ymin=-12, ymax=-2), fill = "orange")+
    geom_rect(aes(x = NULL, y = NULL, xmin=1.742293, xmax=1.743649, ymin=-12, ymax=-2), fill = "orange")+
    # geom_text(aes(x = 0.268191 + (0.961226 - 0.268191) / 2, y = 5,
    #               label = "KHDRBS2"),
    #           size = 5)+
    geom_text(aes(x = 1.448844 + (1.456438 - 1.448844) / 2, y = -15,
                  label = "LOC100141197"),
              size = 5)+
    geom_text(aes(x = 1.742293 + (1.743649 - 1.742293) / 2, y = -15,
                  label = "LOC782044"),
              size = 5)+
    labs(x = "Megabases", y = "", fill = "Epigenetic Mark Type")+
    theme_cowplot()+
    #xlim(c(1,2))+
    ylim(c(17, -17))+
    theme(legend.position = c(0.1, .8),
          axis.text.y = element_text(color = "white"),
          axis.ticks.y = element_line(color = "white"),
          axis.line.y = element_line(color = "white"))


plot_grid(
  plot_grid(CHR_GPSM %>% 
    dplyr::rename(CHR = Chr, BP = bp) %>% 
    filter(p < 0.1) %>% 
    mutate(BP = BP/1000000) %>% 
    ggplot(aes(x = BP, y = -log10(p)))+
    geom_point(alpha = 0.5)+
    geom_point(data = CHR_COJO, aes(x = BP/1000000, y = -log10(p)), color = "indianred", size = 2)+
    scale_x_continuous(breaks = scales::pretty_breaks(n = 5))+
    scale_y_continuous(labels = scales::number_format(accuracy = 0.1))+
    labs(x = "Megabases")+
    #xlim(c(1,2))+
    geom_hline(yintercept = 7.3, color = "red")+
    theme_cowplot(),
  CHR_SGPSM %>% 
  dplyr::rename(CHR = Chr, BP = bp) %>% 
    filter(p<0.1) %>% 
    mutate(BP = BP/1000000) %>% 
    ggplot(aes(x = BP, y = -log10(p)))+
    geom_point(alpha = 0.5)+
    geom_point(data = CHR_SCOJO, aes(x = BP/1000000, y = -log10(p)), color = "indianred", size = 2)+
    scale_x_continuous(breaks = scales::pretty_breaks(n = 5))+
    scale_y_continuous(labels = scales::number_format(accuracy = 0.1))+
    labs(x = "Megabases")+
    #xlim(c(1,2))+
    geom_hline(yintercept = 7.3, color = "red")+
    theme_cowplot(),
  nrow = 1, labels = c("A", "B")),
  CHR_GPSM %>% 
    filter(Chr == 23,
           bp > 1000000,
           bp < 2000000) %>% 
    dplyr::rename(CHR = Chr, BP = bp) %>% 
    mutate(BP = BP/1000000) %>% 
    ggplot(aes(x = BP, y = -log10(p)))+
    geom_point(alpha = 0.5)+
    geom_point(data = CHR_COJO, aes(x = BP/1000000, y = -log10(p)), color = "indianred", size = 2)+
    scale_x_continuous(breaks = scales::pretty_breaks(n = 5))+
    scale_y_continuous(labels = scales::number_format(accuracy = 0.1))+
    labs(x = "")+
    xlim(c(1,2))+
    geom_hline(yintercept = 7.3, color = "red")+
    theme_cowplot(),
  CHR_SGPSM %>% 
    filter(Chr == 23,
           bp > 1000000,
           bp < 2000000) %>% 
    dplyr::rename(CHR = Chr, BP = bp) %>% 
    mutate(BP = BP/1000000) %>% 
    ggplot(aes(x = BP, y = -log10(p)))+
    geom_point(alpha = 0.5)+
    geom_point(data = CHR_SCOJO, aes(x = BP/1000000, y = -log10(p)), color = "indianred", size = 2)+
    scale_x_continuous(breaks = scales::pretty_breaks(n = 5))+
    labs(x = "")+
    xlim(c(1,2))+
    geom_hline(yintercept = 7.3, color = "red")+
    theme_cowplot(),
  functional_plot,
  chr23_af %>% 
  mutate(p = case_when(p == 0 ~ 1e-310,
                        TRUE ~ p)) %>% 
  ggplot(aes(x = birthyear, y = AF, group = rs, color = -log10(p)))+
  geom_smooth(se = FALSE, size = 2)+
  facet_grid(~analysis,)+
  theme_cowplot()+
  ylim(c(0,1))+
  scale_color_viridis_c(option = "inferno", begin = 0.25, end = 0.9)+
  labs(x = "Birth Year", y = "Allele Frequency", color = "-log10(p)")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)),
  nrow = 5,
  labels = c("", "C", "D", "E", "F"),
  rel_heights = c(1,1,1,1,1),
  align = "h"
)

ggsave("output/figures/CHR23.png", width = 12, height = 15,  units = "in")


  
functional_plot
```

## Sequence vs 850K compare

```{r}
CHR_YGPSM = read_tsv("output/200910_RAN/seq_gwas/200910_RAN.age.chr23.mlma.gz")
CHR_YGPSM_CHIP = read_tsv("output/200910_RAN/gwas/200910_RAN.age.850K.mlma.gz") %>% 
  filter(Chr == 23)

plot_grid(
  CHR_YGPSM_CHIP %>% 
      filter(Chr == 23,
             bp > 0,
             bp < 2000000) %>% 
      dplyr::rename(CHR = Chr, BP = bp) %>% 
      mutate(BP = BP/1000000) %>% 
      ggplot(aes(x = BP, y = -log10(p)))+
      geom_point(alpha = 0.5)+
      #geom_point(data = CHR_YCOJO, aes(x = bp/1000000, y = -log10(p)), color = "indianred", size = 2)+
      scale_x_continuous(breaks = scales::pretty_breaks(n = 5))+
      labs(x = "Megabases", title = "850K Imputed")+
      xlim(c(0,2))+
      geom_hline(yintercept = 7.3, color = "red")+
      theme_cowplot(),
  CHR_YGPSM %>% 
      filter(Chr == 23,
             bp > 0,
             bp < 2000000) %>% 
      dplyr::rename(CHR = Chr, BP = bp) %>% 
      mutate(BP = BP/1000000) %>% 
      ggplot(aes(x = BP, y = -log10(p)))+
      geom_point(alpha = 0.5)+
      #geom_point(data = CHR_YCOJO, aes(x = bp/1000000, y = -log10(p)), color = "indianred", size = 2)+
      scale_x_continuous(breaks = scales::pretty_breaks(n = 5))+
      labs(x = "Megabases", title = "Sequence Imputed")+
      xlim(c(0,2))+
      geom_hline(yintercept = 7.3, color = "red")+
      theme_cowplot(),
  nrow = 2
)
```


## Chromosome 2

```{r}
chr2_ldplot = 
  plot_grid(
  read.table("output/200910_RAN/ld/200910_RAN.chr2.young_gpsm_sigsnps.ld.gz", sep="",header=T) %>%  
    filter(SNP_A %in% ran_age_seq_cojo$SNP) %>% 
    mutate(dist = abs(BP_A - BP_B)/1000) %>% 
    filter(R2 > 0.01,
           dist < 100) %>% 
    ggplot(aes(x=dist, y = R2))+
    geom_point(alpha = 0.1,color = "indianred")+
    stat_smooth(method ="loess", se = FALSE, size = 2, alpha = 1, color = "indianred")+
    ylim(c(0,1))+
    theme_cowplot()+
    scale_color_viridis_d()+
    labs(x = "", title = "GPSM", y = expression("r "^2)),
  read.table("output/200910_RAN/ld/200910_RAN.chr2.nSL.ld.gz", sep="",header=T) %>%  
    mutate(dist = abs(BP_A - BP_B)/1000) %>% 
    filter(R2 > 0.01,
           dist < 100) %>% 
    ggplot(aes(x=dist, y = R2))+
    geom_point(alpha = 0.01,color = "springgreen3")+
    stat_smooth(method ="loess", se = FALSE, size = 2, alpha = 1, color = "springgreen3")+
    ylim(c(0,1))+
    theme_cowplot()+
    scale_color_viridis_d()+
    labs(x = "Distance to focal SNP (kb)", y = " ", title = "nSL"),
  read.table("output/200910_RAN/ld/200910_RAN.chr2.RAiSD.ld.gz", sep="",header=T) %>%  
    mutate(dist = abs(BP_A - BP_B)/1000) %>% 
    filter(R2 > 0.01,
           dist < 100) %>% 
    ggplot(aes(x=dist, y = R2))+
    geom_point(alpha = 0.01,color = "slateblue3")+
    stat_smooth(method ="loess", se = FALSE, size = 2, alpha = 1, color = "slateblue3")+
    ylim(c(0,1))+
    theme_cowplot()+
    scale_color_viridis_d()+
    labs(x = "", y = " ", title = "RAiSD"),
  nrow = 1)

chr2_af = 
  bind_rows(
    ran_full_af %>%
      # filter(rs %in% c("2:6757167:A:G", "2:7498211:T:C","2:8100524:G:T", "2:21860593:T:A", "2:27120579:A:G", "2:111403008:C:T", "2:113671290:C:G", "2:127446835:C:G")) %>%
      mutate(analysis = "GPSM (Full)"),
    ran_young_af %>%
    #   filter(rs %in% c("2:7500278:A:G", "2:8100524:G:T", "2:8674849:T:C", "2:9826292:C:T", "2:10679762:C:T", "2:12194425:G:C", "2:12208106:T:C", "2:12363109:A:G", "2:12386721:G:A", "2:12732917:G:A", "2:15449766:T:C", "2:15764320:C:G", "2:16047868:T:C", "2:19635098:C:A", "2:22112854:T:C", "2:22339424:A:G", "2:22745800:G:T", "2:22868999:T:C", "2:23027621:T:C", "2:23259680:G:A", "2:28639397:T:C", "2:28823286:A:T", "2:30374233:T:A", "2:32952841:C:T", "2:42247695:G:A", "2:45247236:T:G", "2:49303806:A:T", "2:53006426:C:G", "2:53126236:G:A", "2:53151542:A:G", "2:53189373:T:G", "2:53283910:C:T", "2:53353546:A:G", "2:53390590:G:A", "2:53730042:G:A", "2:53803504:G:A", "2:60054014:G:A", "2:63289368:G:A", "2:68387467:C:T", "2:76483168:C:T", "2:92863381:G:A", "2:102818589:G:A", "2:103217345:T:A", "2:103225507:A:G", "2:103741954:C:A", "2:104899441:A:G", "2:105543115:G:A", "2:113229961:G:T", "2:136159490:A:G")) %>% 
      mutate(analysis = "GPSM (Young)"),
    read_delim("output/200910_RAN/af_trajectories/200910_RAN.chr2.nsl_afs.raw.gz",
             delim = " ") %>%
  filter(!str_detect(FID, "NA")) %>%
  dplyr::rename(birthyear = FID) %>%
  mutate_at(.vars = vars(contains(":")), .funs = by2) %>%
  select(-PAT, -MAT, -SEX, -PHENOTYPE) %>%
  reshape2::melt(id = c("birthyear", "IID")) %>%
  mutate(rs = str_split_fixed(variable, "_", n = 2)[,1],
         AF = value,
         analysis = "nSL") %>% 
  select(-value, -IID, -variable),
    read_delim("output/200910_RAN/af_trajectories/200910_RAN.chr2.raisd_afs.raw.gz",
             delim = " ") %>%
  filter(!str_detect(FID, "NA")) %>%
  dplyr::rename(birthyear = FID) %>%
  mutate_at(.vars = vars(contains(":")), .funs = by2) %>%
  select(-PAT, -MAT, -SEX, -PHENOTYPE) %>%
  reshape2::melt(id = c("birthyear", "IID")) %>%
  mutate(rs = str_split_fixed(variable, "_", n = 2)[,1],
         AF = value,
         analysis = "RAiSD") %>% 
  select(-value, -IID, -variable)
  )

chr2_afs = 
  chr2_af %>% 
    filter(rs %in% c("2:7498211:T:C", "2:6757167:A:G", "2:21860593:T:A", "2:113671290:C:G", "2:111403008:C:T", "2:53151542:A:G", "2:103217345:T:A", "2:22339424:A:G", "2:7500278:A:G", "2:136159490:A:G", "2:22926713:C:T", "2:16900789:A:C", "2:89172015:A:C", "2:116449514:T:C", "2:5948954:C:T", "2:15116962:G:C", "2:38937843:A:C", "2:44444145:T:C", "2:81303175:C:T", "2:111108096:T:C")) %>% 
    ggplot(aes(x = birthyear, y = AF, group = rs, color = analysis))+
    geom_smooth(se = FALSE)+
    facet_wrap(~analysis, nrow = 1, scales = "free_y")+
    theme_cowplot()+
    ylim(c(0,0.5))+
    labs(x = "Year", y = "Allele Frequency")+
    scale_color_manual(values = c("dodgerblue", "indianred", "springgreen 3", "slateblue3"))+
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.background = element_blank(),
          strip.text = element_text(hjust = 0, size = rel(1.1), face = "bold"))

chr2_afs
```


```{r}
CHR_YGPSM = read_tsv("output/200910_RAN/seq_gwas/200910_RAN.young_age.chr2.mlma.gz")
CHR_YCOJO = filter(ran_young_seq_cojo, CHR == 2)

CHR_GPSM = read_tsv("output/200910_RAN/seq_gwas/200910_RAN.age.chr2.mlma.gz")
CHR_COJO = filter(ran_age_seq_cojo, CHR == 2)
CHR_RAISD = read_tsv("output/200910_RAN/sfs_selection/RAiSD_Report.200910_RAN.chr2.gz",
                 skip = 1,
                 col_names = c("BP", "stat", "finish", "VAR", "SFS", "LD", "Mu"))
CHR_NSL = 
  ran_nsl %>% 
  mutate(CHR = as.numeric(str_split_fixed(window, ":", n = 2)[,1]),
         BP = as.numeric(str_split_fixed(window, ":", n = 2)[,2])) %>% 
  filter(CHR == 2)

CHR_YSIG = 
  CHR_YGPSM %>% 
  filter(p <  4.289e-9) %>% 
  .$SNP

CHR_SIG = 
  CHR_GPSM %>% 
  filter(p <  4.289e-9) %>% 
  .$SNP

chr2_manhattans = 
  plot_grid(
    CHR_GPSM %>% 
      dplyr::rename(CHR = Chr, BP = bp) %>% 
      filter(p < 0.1) %>% 
      mutate(BP = BP/1000000) %>% 
      ggplot(aes(x = BP, y = -log10(p)))+
      geom_point(alpha = 0.5)+
      geom_point(data = CHR_COJO, aes(x = BP/1000000, y = -log10(p)), color = "dodgerblue")+
      #geom_point(data = filter(CHR_GPSM, SNP %in% CHR_YSIG), aes(x = bp/1000000, y = -log10(p)), color = "springgreen", alpha = 0.5)+
      scale_x_continuous(breaks = scales::pretty_breaks(n = 5))+
      labs(x = "Megabases")+
      ylim(c(0,40))+
      geom_hline(yintercept = 7.3, color = "red")+
      theme_cowplot(),
    CHR_YGPSM %>% 
      dplyr::rename(CHR = Chr, BP = bp) %>% 
      filter(p < 0.1) %>% 
      mutate(BP = BP/1000000) %>% 
      ggplot(aes(x = BP, y = -log10(p)))+
      geom_point(alpha = 0.5)+
      geom_point(data = CHR_YCOJO, aes(x = BP/1000000, y = -log10(p)), color = "indianred")+
      #geom_point(data = filter(CHR_YGPSM, SNP %in% CHR_SIG), aes(x = bp/1000000, y = -log10(p)), color = "indianred", alpha = 0.5)+
      scale_x_continuous(breaks = scales::pretty_breaks(n = 5))+
      labs(x = "Megabases")+
      geom_hline(yintercept = 7.3, color = "red")+
      ylim(c(0,40))+
      theme_cowplot(),
    CHR_NSL %>%
      mutate(BP = BP/1000000) %>%
      ggplot(aes(x = BP, y = MeanY))+
      geom_point(alpha = 0.5)+
      scale_x_continuous(breaks = scales::pretty_breaks(n = 5))+
      labs(x = "Megabases", y = "Mean Window |nSL|")+
      geom_hline(yintercept = 2.15, color = "red")+
      theme_cowplot()+
      geom_vline(xintercept = CHR_YCOJO$BP/1000000,
                 color = "indianred",
                 alpha = 0.5)+
      geom_vline(xintercept = CHR_COJO$BP/1000000,
                 color = "dodgerblue",
                 alpha = 0.5),
    CHR_RAISD %>%
      filter(Mu > 10) %>%
      mutate(BP = BP/1000000) %>%
      ggplot(aes(x = BP, y = Mu))+
      geom_point(alpha = 0.5)+
      #geom_point(data = filter(CHR_YGPSM, SNP %in% CHR_SIG), aes(x = bp/1000000, y = -log10(p)), color = "indianred", alpha = 0.5)+
      scale_x_continuous(breaks = scales::pretty_breaks(n = 5))+
      labs(x = "Megabases", y = expression(mu))+
      geom_hline(yintercept = 285, color = "red")+
      theme_cowplot()+
      geom_vline(xintercept = CHR_YCOJO$BP/1000000,
                 color = "indianred",
                 alpha = 0.5)+
      geom_vline(xintercept = CHR_COJO$BP/1000000,
                 color = "dodgerblue",
                 alpha = 0.5),
    nrow = 2,
    align = "v",
    labels = c("A", "B", "C", "D")
  )



plot_grid(chr2_manhattans,
          chr2_ldplot,
          chr2_afs,
          labels = c("", "E", "F"),
          rel_heights = c(1, 0.5, 0.5),
          nrow = 3)

ggsave("output/figures/CHR2.png", width = 12, height = 12, units = "in")
```
### Age Distributions
```{r}
plot_grid(
  redangus %>% 
    select(birth_date) %>%
    ggplot()+
    geom_histogram(aes(x = birth_date), mutate(redangus, z = FALSE), bins = 100, fill = "indianred")+
    geom_histogram(aes(x = birth_date), mutate(redangus, z = TRUE), bins = 250, fill = "indianred")+
    theme_cowplot()+
    labs(x = "Birth Date", y = "Count")+
    facet_zoom(xlim = c(as.Date("1975-01-01"),as.Date("2012-01-01")), ylim = c(0,175), zoom.data = z, horizontal = FALSE),
  simmental %>% 
    select(birth_date) %>%
    ggplot()+
    geom_histogram(aes(x = birth_date), mutate(simmental, z = FALSE), bins = 100, fill = "dodgerblue")+
    geom_histogram(aes(x = birth_date), mutate(simmental, z = TRUE), bins = 250, fill = "dodgerblue")+
    theme_cowplot()+
    labs(x = "Birth Date", y = "Count")+
    facet_zoom(xlim = c(as.Date("1965-01-01"),as.Date("2008-01-01")), ylim = c(0,175), zoom.data = z, horizontal = FALSE), 
  nrow = 2, labels = "AUTO")
ggsave("output/figures/age_distribution.png", width = 10, height = 10, units = "in")
```

### GREML Residual qq-plots
```{r}
make_qq <- function(dd, x) {
    dd<-dd[order(dd[[x]]), ]
    dd$qq <- qnorm(ppoints(nrow(dd)))
    dd
}
plot_grid(
  read_blp("output/200910_RAN/greml/200910_RAN.age.850K.indi.blp") %>% 
      left_join(redangus %>% 
                  select(international_id, age)) %>% 
    make_qq("Residual") %>%
    ggplot()+
    geom_point(aes(x = qq, y = Residual, color = age))+
    labs(x = "Theoretical Res.", y = "Observed Res.", color = "Years since Birth\non 01/01/2020")+
    stat_qq_line(aes(sample = Residual),color = "red")+
    scale_color_viridis_c()+
    theme_cowplot(),
  read_blp("output/200910_RAN/greml/200910_RAN.age.850K.indi.blp") %>% 
    left_join(redangus %>% 
                select(international_id, age)) %>% 
  make_qq("BV") %>%
  ggplot()+
  geom_point(aes(x = qq, y = BV, color = age))+
  labs(x = "Theoretical BV", y = "Observed BV", color = "Years since Birth\non 01/01/2020")+
  stat_qq_line(aes(sample = BV),color = "red")+
  scale_color_viridis_c()+
  theme_cowplot(),
  read_blp("output/200907_SIM/greml/200907_SIM.age.850K.indi.blp") %>% 
    left_join(simmental %>% 
                select(international_id, age)) %>% 
  make_qq("Residual") %>%
  ggplot()+
  geom_point(aes(x = qq, y = Residual, color = age))+
  labs(x = "Theoretical Res.", y = "Observed Res.", color = "Years since Birth\non 01/01/2020")+
  stat_qq_line(aes(sample = Residual),color = "red")+
  scale_color_viridis_c()+
  theme_cowplot(),
  read_blp("output/200907_SIM/greml/200907_SIM.age.850K.indi.blp") %>% 
    left_join(simmental %>% 
                select(international_id, age)) %>% 
  make_qq("BV") %>%
  ggplot()+
  geom_point(aes(x = qq, y = BV, color = age))+
  labs(x = "Theoretical BV", y = "Observed BV", color = "Years since Birth\non 01/01/2020")+
  stat_qq_line(aes(sample = BV),color = "red")+
  scale_color_viridis_c()+
  theme_cowplot(),
  nrow = 2,
  labels = "AUTO")
ggsave("output/figures/raw_age_greml_qq.png", width = 10, height = 8, units = "in")
```

### GREML Residual qq-plots for Box-Cox Transform
```{r}
make_qq <- function(dd, x) {
    dd<-dd[order(dd[[x]]), ]
    dd$qq <- qnorm(ppoints(nrow(dd)))
    dd
}
plot_grid(
  read_blp("output/200910_RAN/greml/200910_RAN.bc_age.850K.indi.blp") %>% 
      left_join(redangus %>% 
                  select(international_id, age)) %>% 
    make_qq("Residual") %>%
    ggplot()+
    geom_point(aes(x = qq, y = Residual, color = age))+
    labs(x = "Theoretical Res.", y = "Observed Res.", color = "Years since Birth\non 01/01/2020")+
    stat_qq_line(aes(sample = Residual),color = "red")+
    scale_color_viridis_c()+
    theme_cowplot(),
  read_blp("output/200910_RAN/greml/200910_RAN.bc_age.850K.indi.blp") %>% 
    left_join(redangus %>% 
                select(international_id, age)) %>% 
  make_qq("BV") %>%
  ggplot()+
  geom_point(aes(x = qq, y = BV, color = age))+
  labs(x = "Theoretical BV", y = "Observed BV", color = "Years since Birth\non 01/01/2020")+
  stat_qq_line(aes(sample = BV),color = "red")+
  scale_color_viridis_c()+
  theme_cowplot(),
  read_blp("output/200907_SIM/greml/200907_SIM.bc_age.850K.indi.blp") %>% 
    left_join(simmental %>% 
                select(international_id, age)) %>% 
  make_qq("Residual") %>%
  ggplot()+
  geom_point(aes(x = qq, y = Residual, color = age))+
  labs(x = "Theoretical Res.", y = "Observed Res.", color = "Years since Birth\non 01/01/2020")+
  stat_qq_line(aes(sample = Residual),color = "red")+
  scale_color_viridis_c()+
  theme_cowplot(),
  read_blp("output/200907_SIM/greml/200907_SIM.bc_age.850K.indi.blp") %>% 
    left_join(simmental %>% 
                select(international_id, age)) %>% 
  make_qq("BV") %>%
  ggplot()+
  geom_point(aes(x = qq, y = BV, color = age))+
  labs(x = "Theoretical BV", y = "Observed BV", color = "Years since Birth\non 01/01/2020")+
  stat_qq_line(aes(sample = BV),color = "red")+
  scale_color_viridis_c()+
  theme_cowplot(),
  nrow = 2,
  labels = "AUTO")
ggsave("output/figures/bc_age_greml_qq.png", width = 10, height = 8, units = "in")
```

### Log ransformation qq-plots GREML
```{r}
plot_grid(
  read_blp("output/200910_RAN/greml/200910_RAN.nc_age.850K.indi.blp") %>% 
      left_join(redangus %>% 
                  select(international_id, age)) %>% 
    make_qq("Residual") %>%
    ggplot()+
    geom_point(aes(x = qq, y = Residual, color = age))+
    labs(x = "Theoretical", y = "Observed", color = "Years since Birth\non 01/01/2020")+
    stat_qq_line(aes(sample = Residual),color = "red")+
    scale_color_viridis_c()+
    theme_cowplot(),
  read_blp("output/200910_RAN/greml/200910_RAN.log_age.850K.indi.blp") %>% 
    left_join(redangus %>% 
                select(international_id, age)) %>% 
  make_qq("BV") %>%
  ggplot()+
  geom_point(aes(x = qq, y = BV, color = age))+
  labs(x = "Theoretical", y = "Observed", color = "Years since Birth\non 01/01/2020")+
  stat_qq_line(aes(sample = BV),color = "red")+
  scale_color_viridis_c()+
  theme_cowplot(),
  read_blp("output/200907_SIM/greml/200907_SIM.log_age.850K.indi.blp") %>% 
    left_join(simmental %>% 
                select(international_id, age)) %>% 
  make_qq("Residual") %>%
  ggplot()+
  geom_point(aes(x = qq, y = Residual, color = age))+
  labs(x = "Theoretical", y = "Observed", color = "Years since Birth\non 01/01/2020")+
  stat_qq_line(aes(sample = Residual),color = "red")+
  scale_color_viridis_c()+
  theme_cowplot(),
  read_blp("output/200907_SIM/greml/200907_SIM.log_age.850K.indi.blp") %>% 
    left_join(simmental %>% 
                select(international_id, age)) %>% 
  make_qq("BV") %>%
  ggplot()+
  geom_point(aes(x = qq, y = BV, color = age))+
  labs(x = "Theoretical", y = "Observed", color = "Years since Birth\non 01/01/2020")+
  stat_qq_line(aes(sample = BV),color = "red")+
  scale_color_viridis_c()+
  theme_cowplot(),
  nrow = 2,
  labels = "AUTO")
ggsave("output/figures/log_age_greml_qq.png", width = 10, height = 8, units = "in")
```
### Transform Manhattans Red Angus
```{r}
ran_gpsm_age = 
 read_gwas("output/200910_RAN/gwas/200910_RAN.age.850K.mlma.gz")
ran_gpsm_youngage = 
 read_gwas("output/200910_RAN/gwas/200910_RAN.young_age.850K.mlma.gz")
ran_gpsm_logage = 
 read_gwas("output/200910_RAN/gwas/200910_RAN.log_age.850K.mlma.gz")

ssnps = filter(ran_gpsm_logage, p < 6.17e-8) %>% .$SNP

plot_grid(
  ggmanhattan2(ran_gpsm_age,
               prune = 0.01, 
               sig_threshold_p = 6.17e-8,
               sigsnps = ssnps),
  ggmanhattan2(ran_gpsm_age,
               prune = 0.01, 
               sig_threshold_p = 6.17e-8,
               sigsnps = ssnps)+
    ylim(c(0,15)),
  ggmanhattan2(ran_gpsm_logage,
               prune = 0.01, 
               sig_threshold_p = 6.17e-8),
  ggmanhattan2(ran_gpsm_logage,
               prune = 0.01, 
               sig_threshold_p = 6.17e-8)+
    ylim(c(0,15)),
  ggmanhattan2(ran_gpsm_youngage,
               prune = 0.01, 
               sig_threshold_p = 6.17e-8,
               sigsnps = ssnps),
  ggmanhattan2(ran_gpsm_youngage,
               prune = 0.01, 
               sig_threshold_p = 6.17e-8,
               sigsnps = ssnps)+
    ylim(c(0,15)),
  nrow = 3,
  labels = "AUTO")
ggsave("output/figures/transformation_manhattans.png", width = 10, height = 8, units = "in")
```


### Sweep Manhattan Plots
```{r}

plot_grid(
  sim_nsl %>% dplyr::rename(SNP = window) %>% 
    filter(SNPcount > 3) %>% 
    selscan_manhattans(col = MeanY)+
    geom_hline(yintercept = 2.38, color = "red")+
    labs(y = "Mean nSL value"),
  sim_raisd%>%
    selscan_manhattans(col = Mu)+
    geom_hline(yintercept = 193.5, color = "red")+
    labs(y = expression(paste("RAiSD ", mu, " value"))),
nrow = 2, align = "h")

ggsave("output/figures/SIM.sweep.png", width = 12, height = 5, units = "in")

plot_grid(
  ran_nsl %>% dplyr::rename(SNP = window) %>% 
    filter(SNPcount > 3) %>% 
    selscan_manhattans(col = MeanY)+
    geom_hline(yintercept = 2.34, color = "red")+
    labs(y = "Mean nSL value"),
  ran_raisd%>%
    selscan_manhattans(col = Mu)+
    geom_hline(yintercept = 285, color = "red")+
    labs(y = expression(paste("RAiSD ", mu, " value"))),
nrow = 2, align = "h")

ggsave("output/figures/RAN.sweep.png", width = 12, height = 5, units = "in")

```

### QQ Plots

```{r}

plot_grid(
  read_tsv("output/200910_RAN/seq_gwas/200910_RAN.full_age.pvalues.txt.gz",
           col_names = c("p"),
           col_types = cols(p = col_double())) %>% 
    .$p %>% 
    ggqq(),
  read_tsv("output/200910_RAN/seq_gwas/200910_RAN.young_age.pvalues.txt.gz",
           col_names = c("p"),
           col_types = cols(p = col_double())) %>% 
    .$p %>% 
    ggqq(),
  read_tsv("output/200907_SIM/seq_gwas/200907_SIM.full_age.pvalues.txt.gz",
           col_names = c("p"),
           col_types = cols(p = col_double())) %>% 
    #filter(p < 0.1) %>% 
    .$p %>% 
    ggqq(),
  read_tsv("output/200907_SIM/seq_gwas/200907_SIM.pb_age.pvalues.txt.gz",
           col_names = c("p"),
           col_types = cols(p = col_double())) %>%
    .$p %>% 
    ggqq(),
  nrow = 2,
  labels = "AUTO"
  )

ggsave("output/figures/QQPLOTS.png", height = 10, width = 10, units = "in")
```

### WCGALP GPSM FIGURE
```{r}
young_ran_manhattan= 
ran_seq_gpsm_young_age %>% 
    ggmanhattan2(inputfile = ., value = p, 
                 sig_threshold_p = 4.289e-9,
                 sigsnps = ran_young_seq_cojo$SNP)+
    geom_hline(yintercept = 7.3, color = "blue")+
    ylim(1, max(c(-log10(ran_seq_gpsm_age$p), -log10(ran_seq_gpsm_young_age$p))))

full_ran_manhattan= 
ran_seq_gpsm_age %>% 
    ggmanhattan2(inputfile = ., value = p, 
                 sig_threshold_p = 4.289e-9,
                 sigsnps = ran_age_seq_cojo$SNP)+
    geom_hline(yintercept = 7.3, color = "blue")+
    ylim(1, max(c(-log10(ran_seq_gpsm_age$p), -log10(ran_seq_gpsm_young_age$p))))

full_sim_manhattan =
  sim_seq_gpsm_age %>% 
    ggmanhattan2(sig_threshold_p = 4.289e-9,
                 sigsnps2 = sim_seq_cojo$SNP)+
    geom_hline(yintercept = 7.3, color = "blue", size = 0.25)+
    ylim(1, max(c(-log10(sim_seq_gpsm_purebred_age$p), -log10(sim_seq_gpsm_age$p))))

pb_sim_manhattan =
  sim_seq_gpsm_purebred_age %>% 
    ggmanhattan2(sig_threshold_p = 4.289e-9,
                 sigsnps2 = sim_pb_seq_cojo$SNP)+
    geom_hline(yintercept = 7.3, color = "blue", size = 0.25)+
    ylim(1, max(c(-log10(sim_seq_gpsm_purebred_age$p), -log10(sim_seq_gpsm_age$p))))


plot_grid(full_ran_manhattan,
          full_sim_manhattan,
          young_ran_manhattan,
          pb_sim_manhattan,
          nrow = 2, labels = c("A", "C", "B", "D"))

ggsave("output/figures/WCGALP.png", width = 10, height = 4, units = "in")

max(c(-log10(sim_seq_gpsm_purebred_age$p), -log10(sim_seq_gpsm_age$p)))
```

