---
title: "Map Creation"
author: "Troy Rowan"
date: "2020-07-09"
output: 
  html_document:
    code_folding: show
editor_options:
  chunk_output_type: inline
---


```{r, include=FALSE}
set.seed(45288)
knitr::opts_chunk$set(echo = TRUE)
library(workflowr)
library(knitr)
library(rgdal)
library(raster)
library(fpc)
library(RStoolbox)
library(maps)
library(ggthemes)
library(cowplot)
library(purrr)
library(factoextra)
source(here::here("code", "map_functions.R"))
library(corrr)
library(ggcorrplot)
library(factoextra)
library(here)
library(tidyverse)
setwd(here())
```


## Introduction
This uses k-means clustering and 30-year normal climate variables to divide the United States into distinct ecoregions.



## Reading in climate raster files

These originate from the [Prism Climate Group's](https://prism.oregonstate.edu/normals/) website in .bil format 
I was having issues with the rgdal package reading these or any other form, so had to read them and save as RDS files on Workstation (the only place where rgdal installs properly). SCP'd RDS files here, and these read in the individal environment variable rasters. 
```{r, eval=TRUE, warning=FALSE}
temp_raster = 
  readRDS(
    here::here("data", "prism_climate_data", "temp_raster.RDS"))

precip_raster = 
  readRDS(
    here::here("data", "prism_climate_data", "precip_raster.RDS"))

elev_raster = 
  readRDS(
    here::here("data", "prism_climate_data", "elev_raster.RDS"))

dewpt_raster = 
  readRDS(
    here::here("data", "prism_climate_data", "mean_dwpt_raster.RDS")
  )

min_vap_raster = 
  readRDS(
    here::here("data", "prism_climate_data", "min_vpd_raster.RDS"))

max_vap_raster = 
  readRDS(
    here::here("data", "prism_climate_data", "max_vpd_raster.RDS"))

min_temp_raster = 
  readRDS(
    here::here("data", "prism_climate_data", "min_temp_raster.RDS"))

max_temp_raster = 
  readRDS(
    here::here("data", "prism_climate_data", "max_temp_raster.RDS"))
```

```{r, eval=TRUE, echo=TRUE, warning=FALSE}
stacked_raster = 
  stack(
    temp_raster,
    precip_raster,
    elev_raster,
    dewpt_raster,
    max_temp_raster,
    min_temp_raster,
    max_vap_raster,
    min_vap_raster
  ) 

threevar_stacked_raster = 
  stack(
    temp_raster,
    precip_raster,
    elev_raster
  )

fourvar_stacked_raster = 
  stack(
    temp_raster,
    precip_raster,
    elev_raster,
    min_vap_raster
  )

```

## Environmental varible correlations

### Numerical Correlations for complete environmental data 
```{r}
env_correlations = 
  stacked_raster %>% 
    as.data.frame() %>% 
    rename("Temperature" = "band1.1",
      "Precipitation" = "band1.2",
      "Elevation" = "band1.3",
      "DewPoint" = "band1.4",
      "MaxTemperature" = "band1.5",
      "MinTemperature" = "band1.6",
      "MaxVaporPressure" = "band1.7",
      "MinVaporPressure" = "band1.8") %>% 
    na.omit() %>% 
    cor()

env_correlations_p_values = 
  stacked_raster %>% 
    as.data.frame() %>% 
    rename("Temperature" = "band1.1",
      "Precipitation" = "band1.2",
      "Elevation" = "band1.3",
      "DewPoint" = "band1.4",
      "MaxTemperature" = "band1.5",
      "MinTemperature" = "band1.6",
      "MaxVaporPressure" = "band1.7",
      "MinVaporPressure" = "band1.8") %>% 
    na.omit() %>% 
  cor_pmat()

kable(round(env_correlations, 3))
```


### Correlation heat plots
This plot uses heirchical clustering to group variables


```{r}
ggcorrplot(env_correlations,
           hc.order = TRUE)
```

## PCA of all PRISM environmental variables {.tabset}

```{r, eval=TRUE, echo=FALSE}
env_pca =
  stacked_raster %>% 
    as.data.frame() %>%
    na.omit() %>% 
    rename("Temperature" = "band1.1",
      "Precipitation" = "band1.2",
      "Elevation" = "band1.3",
      "DewPoint" = "band1.4",
      "MaxTemperature" = "band1.5",
      "MinTemperature" = "band1.6",
      "MaxVaporPressure" = "band1.7",
      "MinVaporPressure" = "band1.8") %>%
    prcomp(scale = TRUE)
```

### PCA Plot (PC 1 & PC 2)
This shows only loadings of variables (thought that 481,630 was too many points)
```{r}
fviz_pca_var(env_pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )+
  ggtitle("Environmental Variable PCA (PC1 & PC2")
```

### Scree Plot
Vast majority of environmental variance is explained by first two PCs

```{r}
fviz_eig(env_pca)
```




## Identifying best K for full- and three-variable datasets

Using fpc::pamk(), K between 3 and 10
This doesn't run, but appears that based on the fviz_nbclust() function in that the appropriate 
 
It appears that k=2 is optimal by these metrics, but for some reason this step exceeds memory necessary to add to website. While K = 2 may be optimal in a statistical sense, it doesn't necessarily mean that it's useful for our analysis (being able to group individuals into subpopulations)
```{r, eval=FALSE, warning=FALSE}
stacked_raster %>% 
  as.data.frame() %>% 
  na.omit() %>% 
  sample_n(20000) %>% 
  fviz_nbclust(FUNcluster = kmeans)+
  ggtitle("Full Environmental Data")
  #pamk(krange = 3:20, criterion="multiasw", usepam=FALSE, scaling=TRUE)

threevar_stacked_raster %>% 
  as.data.frame() %>% 
  na.omit() %>% 
  sample_n(20000) %>% 
  fviz_nbclust(FUNcluster = kmeans)+
  ggtitle("Three Environmental Variables ")
  #pamk(krange = 3:20, criterion="multiasw", usepam=FALSE, scaling=TRUE)  

fourvar_stacked_raster %>% 
  as.data.frame() %>% 
  na.omit() %>% 
  sample_n(20000) %>% 
  fviz_nbclust(FUNcluster = kmeans)+
  ggtitle("Three Environmental Variables ") %>% 
  #pamk(krange = 3:20, criterion="multiasw", usepam=FALSE, scaling=TRUE)
```

## K-means clustering climate maps {.tabset}
These are the resulting maps from using the full PRISM dataset (all 8 variables) or just three (Temperature, Precipitation, Elevation).

Here, since this is a completely stochastic process, colors don't have any meaning between K's or between all data vs. three variables. 

I'm particulary intersted in how K=10 breaks out the "Fescue Belt" in both the ALL and Three variable cases. 
```{r K-means Maps, fig.height=4, fig.width=12, warning=FALSE}

# plotlist =
#   seq(2,10) %>%
#   purrr::map(
#   ~plot_grid(
#     unsuperClass(stacked_raster,
#                nSamples=2000,
#                nClasses = .x,
#                norm=TRUE,
#                nStarts=5,
#                clusterMap=FALSE) %>%
#     .$map %>%
#     as.data.frame(xy=TRUE) %>%
#     mutate(layer = as.factor(layer)) %>%
#     kmeans_map()+
#       ggtitle(paste("All Variables, k =", .x)),
#     unsuperClass(threevar_stacked_raster,
#                  nSamples=1000,
#                  nClasses = .x,
#                  norm=TRUE,
#                  nStarts=5,
#                  clusterMap=FALSE) %>%
#     .$map %>%
#     as.data.frame(xy=TRUE) %>%
#     mutate(layer = as.factor(layer)) %>%
#     kmeans_map()+
#       ggtitle(paste("Three Variables, k =", .x))
#     )
#   )
# 
# saveRDS(plotlist, "../output/kmeans_plotlist.RDS")


plotlist = readRDS(here::here("output", "kmeans_plotlist.RDS"))
# 
# for (i in 1:length(plotlist)){
#   cat("### ", i)
#   print(plotlist[[i]])
#   #cat(' \n\n')
# }



  
```

### K = 2

```{r, fig.height=4, fig.width=12, warning=FALSE}
plotlist[[1]]
```

### K = 3

```{r, fig.height=4, fig.width=12, warning=FALSE}
plotlist[[2]]
```

### K = 4

```{r, fig.height=4, fig.width=12, warning=FALSE}
plotlist[[3]]
```

### K = 5

```{r, fig.height=4, fig.width=12, warning=FALSE}
plotlist[[4]]
```

### K = 6

```{r, fig.height=4, fig.width=12, warning=FALSE}
plotlist[[5]]
```

### K = 7

```{r, fig.height=4, fig.width=12, warning=FALSE}
plotlist[[6]]
```

### K = 8

```{r, fig.height=4, fig.width=12, warning=FALSE}
plotlist[[7]]
```

### K = 9

```{r, fig.height=4, fig.width=12, warning=FALSE}
plotlist[[8]]
```


### K = 10

```{r, fig.height=4, fig.width=12, warning=FALSE}
plotlist[[9]]
```


## K = 9 ecoregion changes 

I'd noticed eariler that there were major changes that occurred in the K = 9 map when using a different seed. In some cases, when using the these look similar to the original 3-variable ecoregions, and occasionally they break out the Fescue Belt into an extra region. 

NOTE: The colors here aren't corresponding from map to map because "zone number" is stochastically determined from run to run 

Here, generate a k-means clustering map with K = 9 four times. I use each combination of all environmental variables and the three main variables and two different random seeds. 

Using all variables actually makes more sense to me. Correlated variables don't necessarily drive the assignments one way or another. 

Observations:
* More accurate "Fescue Belt" in ALL-Seed 1, that said, still likely too far West, not far enough North, disappears with Seed 2
* Rainforest ecoregion disappears with ALL-Seed 1 map (which is probably okay)
* All-variable maps division of Dakotas tracks Missouri River (and a real chang in production environment)
* ALL-Seed 1 is much less "granular" in elevation divisions. Should result in less weird "regional islands" due to (usually) elevation being created.
* Not sure about the "rain line" in Texas, it certainly gets shifted to the West a bit
* California is a bit strange in all of these maps still, but I think the differences that it picks up are real. 

NOTE: There are definitely issues with coloring in the K = 10 for some reason.

*My suggestion moving forward* 
Use all-variable map (seed #1 version). I think that it makese divisions closer to how I'd divide the US based on what I know about these production environments. I'm interested in potentially partitioning out the Fescue Belt and Southeast further. 
```{r, fig.height=8, fig.width=13, warning=FALSE}
set.seed(325333)
k9_all =
  unsuperClass(stacked_raster,
               nSamples=2000,
               nClasses = 9,
               norm=TRUE,
               nStarts=5,
               clusterMap=FALSE) %>%
    .$map %>%
    as.data.frame(xy=TRUE) %>%
    mutate(layer = as.factor(layer)) %>% 
    filter(!is.na(layer)) %>% 
  mutate(region = case_when(layer == 1 ~ "HighPlains",
                            layer == 2 ~ "Desert",
                            layer == 3 ~ "Southeast",
                            layer == 4 ~ "CornBelt",
                            layer == 5 ~ "FescueBelt",
                            layer == 6 ~ "ForestedMountains",
                            layer == 7 ~ "UpperMidwest",
                            layer == 8 ~ "AridPrairie",
                            layer == 9 ~ "ExtremeDesert"))
#saveRDS(k9_all, "output/k9.allvars.seed1.rds")

k9_three =
  unsuperClass(threevar_stacked_raster,
               nSamples=2000,
               nClasses = 9,
               norm=TRUE,
               nStarts=5,
               clusterMap=FALSE) %>%
    .$map %>%
    as.data.frame(xy=TRUE) %>%
    mutate(layer = as.factor(layer)) %>% 
    filter(!is.na(layer)) %>% 
  mutate(region = case_when(layer == 7 ~ "HighPlains",
                            layer == 5 ~ "Desert",
                            layer == 2 ~ "Southeast",
                            layer == 8 ~ "Rainforest",
                            layer == 1 ~ "FescueBelt",
                            layer == 9 ~ "ForestedMountains",
                            layer == 6 ~ "UpperMidwest",
                            layer == 4 ~ "AridPrairie",
                            layer == 3 ~ "Foothills"))
#saveRDS(k9_three, "output/k9.threevars.seed1.rds")


set.seed(8675305)
k9_all_newseed =
  unsuperClass(stacked_raster,
               nSamples=2000,
               nClasses = 9,
               norm=TRUE,
               nStarts=5,
               clusterMap=FALSE) %>%
    .$map %>%
    as.data.frame(xy=TRUE) %>%
    mutate(layer = as.factor(layer)) %>% 
    filter(!is.na(layer))%>% 
  mutate(region = case_when(layer == 1 ~ "HighPlains",
                            layer == 5 ~ "Desert",
                            layer == 9 ~ "Southeast",
                            layer == 4 ~ "Rainforest",
                            layer == 7 ~ "FescueBelt",
                            layer == 2 ~ "ForestedMountains",
                            layer == 3 ~ "UpperMidwest",
                            layer == 8 ~ "AridPrairie",
                            layer == 3 ~ "ExtremeDesert"))
#saveRDS(k9_all_newseed, "output/k9.allvars.seed2.rds")  

k9_three_newseed =
  unsuperClass(threevar_stacked_raster,
               nSamples=2000,
               nClasses = 9,
               norm=TRUE,
               nStarts=5,
               clusterMap=FALSE) %>%
    .$map %>%
    as.data.frame(xy=TRUE) %>%
    mutate(layer = as.factor(layer)) %>% 
    filter(!is.na(layer))
saveRDS(k9_three_newseed, "output/k9.threevars.seed2.rds")


set.seed(325333)
k9_four =
  unsuperClass(fourvar_stacked_raster,
               nSamples=2000,
               nClasses = 9,
               norm=TRUE,
               nStarts=5,
               clusterMap=FALSE) %>%
    .$map %>%
    as.data.frame(xy=TRUE) %>%
    mutate(layer = as.factor(layer)) %>% 
    filter(!is.na(layer)) %>%
    mutate(region = case_when(layer == 1 ~ "HighPlains", 
                              layer == 2 ~ "Foothills", 
                              layer == 3 ~ "Southeast", 
                              layer == 4 ~ "FescueBelt",
                              layer == 5 ~ "Rainforest",
                              layer == 6 ~ "ForestedMountains",
                              layer == 7 ~ "UpperMidwest",
                              layer == 8 ~ "AridPrairie",
                              layer == 9 ~ "Desert"))



set.seed(325333)
k10 =
  unsuperClass(threevar_stacked_raster,
               nSamples=2000,
               nClasses = 10,
               norm=TRUE,
               nStarts=5,
               clusterMap=FALSE) %>%
    .$map %>%
    as.data.frame(xy=TRUE) %>%
    mutate(layer = as.factor(layer)) %>% 
    filter(!is.na(layer))%>% 
    mutate(region = case_when(layer == 1 ~ "HighPlains",
                              layer == 2 ~ "Desert",
                              layer == 3 ~ "Southeast",
                              layer == 4 ~ "CornBelt",
                              layer == 5 ~ "FescueBelt",
                              layer == 6 ~ "ForestedMountains",
                              layer == 6 ~ "UpperMidwest",
                              layer == 8 ~ "Rainforest",
                              layer == 9 ~ "ExtremeDesert",
                              layer == 10 ~ "Foothills"))

#saveRDS(k10, "output/k10.allvars.seed2.rds")

colorkey = c("HighPlains"="springgreen3", "UpperMidwest"="slateblue2", "Desert"="tomato2", "AridPrairie"="goldenrod1", "FescueBelt"="gray17", "Rainforest" = "brown", "CornBelt"="brown", "Southeast"="darkslategray4", "ForestedMountains"="deeppink3", "Foothills" = "gray50", "ExtremeDesert" = "gray50")
colour = c("10" = "white", "3"="springgreen3", "9"="slateblue2", "1"="tomato2", "5"="goldenrod1", "6"="gray50", "8"="gray17", "4"="brown", "2"="darkslategray4", "7"="deeppink3")
    


plot_grid(
  k9_all %>% 
    kmeans_map(id = region, color = colorkey)+
      ggtitle(paste("    All Variables, k = 9, Seed #1")),
  k9_all_newseed %>% 
    kmeans_map(id = region, color = colorkey)+
      ggtitle(paste("    All Variables, k = 9, Seed #2")),  
  k9_three %>% 
    kmeans_map(id = region, color = colorkey)+
      ggtitle(paste("    Three Variables, k = 9, Seed #1")),
  k10 %>% 
    kmeans_map(id = region, color = colorkey)+
      ggtitle(paste("    All Variables, k = 10, Seed #1")),
  nrow = 2
  )
```

## Comparing K=9 with different input variables {.tabset}

### Three Variables
Temperature (mean), Precipitation, Elevation
```{r, fig.height=4, fig.width=7, warning=FALSE}
k9_three %>%
  kmeans_map(id = region, color = colorkey)+
      ggtitle(paste("    Three Variables, k = 9, Seed #1"))

```


### Four Variables 
Temperature (mean), Precipitation, Elevation, Minimum Vapor Pressure
```{r, fig.height=4, fig.width=7, warning=FALSE}
k9_four %>% 
  kmeans_map(id = region, color = colorkey)+
      ggtitle(paste("    Four Variables, k = 9, Seed #1"))
```


### All Variables
Temperature (min, max, mean), Precipitation, DewPoint, Elevation, Vapor Pressure (min, max)
```{r, fig.height=4, fig.width=7, warning=FALSE}
k9_all %>%
  kmeans_map(id = region, color = colorkey)+
      ggtitle(paste("    All Variables, k = 9, Seed #1"))
```

