---
title: "Generation Proxy Phenotype Exploration"
author: "Troy Rowan"
date: "2020-09-16"
output: 
  html_document:
    code_folding: show
editor_options:
  chunk_output_type: inline
---

```{r setup, include = FALSE}
set.seed(325333)
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
knitr::opts_knit$set(root.dir="C:/Users/troyrowan/Documents/GitHub/local_adaptation_sequence/")
library(workflowr)
library(knitr)
library(maps)
library(ggthemes)
library(cowplot)
library(purrr)
library(factoextra)
source(here::here("code", "map_functions.R"))
#library(ggcorrplot)
library(factoextra)
library(here)
library(lubridate)
library(pedigree)
library(ggforce)
#library(MASS)
library(tidyverse)
library(EnvStats)
library(optiSel)
library(car)
#source("code/GCTA_functions.R")
```


# Simmental


##Full Simmental dataset phenotype file creation

```{r}

simmental =
  read_csv("output/200907_SIM/phenotypes/200907_SIM.info.csv") %>%
      mutate(sqrt_age = age^0.5,
             cbrt_age = age^0.333,
             log_age = log(age),
             bc_age = bcPower(age, lambda = 0.0345))




sim_fam = 
  read_delim("data/200907_SIM/200907_SIM.fam",
            delim = " ",
             col_names = c("fam_id", "international_id", "sire", "dam", "sex", "phenotype"))

#Counts generation number of simmentals (by various metrics)
# sim_gens = 
#   read_csv("data/200907_SIM/mizzou-data-request/6-gen-ped.csv")%>% 
#     select(id = animal, sire, dam) %>% 
#     select(Indiv = id, Sire = sire, Dam = dam) %>% 
#     summary.Pedig()

sim_fam %>%
  left_join(filter(simmental, asa_nbr != 2326708)) %>% 
  #left_join(k9_zips, by = c("breeder_zip" = "zip"))  %>% 
  filter(!is.na(breeder_zip)) %>% 
  distinct() %>% 
  left_join(read_csv("data/200907_SIM/mizzou-birthdates.csv") %>% 
              select(international_id = int_id, birth_date = anm_dob)) %>% 
  mutate(birth_year = year(birth_date),
         age = time_length(interval(start = birth_date, end = Sys.Date()), "year")) %>% 
  left_join(sim_gens, by = c("international_id" = "Indiv")) #%>% 
  #write_csv("output/200907_SIM/phenotypes/200907_SIM.info.csv")


simmental = 
  read_csv("output/200907_SIM/phenotypes/200907_SIM.info.csv") %>%
  mutate(ASA_nbr = asa_nbr,
         dummyid = international_id) %>% 
  left_join(read_csv("data/200907_SIM/breeds.csv.gz") %>% 
  pivot_wider(names_from = brd_cd, values_from = percent) %>% 
    replace(is.na(.), 0))
  

```

### Comparing %SIM over time

Looks like an inflection point 
```{r}
simmental %>% 
  ggplot(aes(birth_date, SM))+
  geom_smooth()+
  geom_point(alpha = 0.1)

simmental %>% 
  ggplot(aes(x=birth_year, y = SM*100))+
  geom_count(alpha = 0.7)+
  geom_smooth(color = "red")+
  labs(x = "Birth Year", y = "% Simmental Ancestry")+
  theme_cowplot()
```

Outputting phenotype files
```{r}

#Generates optimal BoxCox Lambda value
#boxcox(simmental$age, optimize = TRUE)

sim_fam %>% 
  left_join(filter(simmental, 
                   asa_nbr != 2326708)) %>% 
  mutate(FID = international_id) %>% 
  select(FID, IID = international_id, age, birth_year, equiGen, fullGen, maxGen, Inbreeding) %>% 
  distinct() %>%
    mutate(sqrt_age = age^0.5,
           cbrt_age = age^0.333,
           log_age = log(age),
           bc_age = bcPower(age, lambda = 0.0345),
           young_age = case_when(birth_year >= 2012 ~ age,
                                 TRUE ~ NA_real_),
           young_log_age = case_when(birth_year >= 2012 ~ log_age,
                                 TRUE ~ NA_real_),
           equiGen_nozero = case_when(equiGen <= 1 ~ NA_real_,
                                 TRUE ~ equiGen),
         old_age = case_when(birth_year < 2012 ~ age,
                               TRUE ~ NA_real_),
         old_log_age = case_when(birth_year < 2012 ~ log_age,
                               TRUE ~ NA_real_))%>% 
  group_by(IID) %>% 
  slice(1:1) %>% 
  ungroup() %>% 
  write_tsv("output/200907_SIM/phenotypes/200907_SIM.generation_proxy.txt", col_names = FALSE) 


sim_fam %>% 
  left_join(filter(simmental, 
                   asa_nbr != 2326708)) %>% 
  mutate(FID = international_id) %>% 
  select(FID, IID = international_id, age, birth_year, equiGen, fullGen, maxGen, Inbreeding, SM, AN) %>% 
  distinct() %>%
    mutate(full_age = case_when(SM > 0.05 ~ age,
                                TRUE ~ NA_real_),
           full_log_age = case_when(SM > 0.05 ~ log(age),
                                    TRUE ~ NA_real_),
           full_young_age = case_when(birth_year > 2008 & SM > 0.05 ~ age,
                                TRUE ~ NA_real_),
           full_old_age = case_when(birth_year <= 2008 & SM > 0.05 ~ age,
                                      TRUE ~ NA_real_),
           simangus3050_age = case_when(SM < 0.30 & AN > 0.50 ~ age,
                                    TRUE ~ NA_real_),
           simangus2070_age = case_when(SM > 0.20 & SM < 0.7 ~ age,
                                    TRUE ~ NA_real_),
           sim70_age = case_when(SM > 0.70 ~ age,
                                TRUE ~ NA_real_),
           sim70_log_age = case_when(SM > 0.70 ~ log(age),
                                    TRUE ~ NA_real_),
           pb_age = case_when(SM > 0.99 ~ age,
                              TRUE ~ NA_real_),
           pb_log_age = case_when(SM > 0.99 ~ log(age),
                              TRUE ~ NA_real_),
           pb_young_age = case_when(birth_year > 2008 & SM > 0.99 ~ age,
                                    TRUE ~ NA_real_),
           pb_old_age = case_when(birth_year <= 2008 & SM > 0.99 ~ log(age),
                              TRUE ~ NA_real_),
           pb_young_age = case_when(birth_year > 2008 & SM > 0.99 ~ log(age),
                                    TRUE ~ NA_real_)
           ) %>% 
  group_by(IID) %>% 
  slice(1:1) %>% 
  ungroup() %>%
  select(-SM, -AN, -age, -birth_year, -equiGen, -fullGen, -maxGen, -Inbreeding) %>% head()
  write_tsv("output/200907_SIM/phenotypes/200907_SIM.breedprop_generation_proxy.txt", col_names = FALSE) 


  
```

##Subsetting Purebred Simmentals
```{r}
sim_fam %>%
  left_join(filter(simmental, asa_nbr != 2326708)) %>%
  mutate(ASA_nbr = asa_nbr,
         dummyid = international_id) %>% 
  left_join(read_csv("data/200907_SIM/breeds.csv") %>% 
  pivot_wider(names_from = brd_cd, values_from = percent)) %>% 
  filter(SM == 1) %>% 
  select(dummyid, international_id) %>% 
  write_tsv("output/201020_PBSIM/201020_PBSIM.keepids.txt", col_names = FALSE)

sim_fam %>%
  left_join(filter(simmental, asa_nbr != 2326708)) %>%
  mutate(ASA_nbr = asa_nbr,
         dummyid = international_id) %>% 
  left_join(read_csv("data/200907_SIM/breeds.csv") %>% 
  pivot_wider(names_from = brd_cd, values_from = percent)) %>% 
  filter(SM == 1) %>%
    mutate(sqrt_age = age^0.5,
           cbrt_age = age^0.333,
           log_age = log(age),
           bc_age = bcPower(age, lambda = 0.0345),
           young_age = case_when(birth_year >= 2012 ~ age,
                                 TRUE ~ NA_real_),
           young_log_age = case_when(birth_year >= 2012 ~ log_age,
                                 TRUE ~ NA_real_),
           equiGen_nozero = case_when(equiGen <= 1 ~ NA_real_,
                                 TRUE ~ equiGen),
         old_age = case_when(birth_year < 2012 ~ age,
                               TRUE ~ NA_real_),
         old_log_age = case_when(birth_year < 2012 ~ log_age,
                               TRUE ~ NA_real_)) %>% 
  select(dummyid, international_id, age, sqrt_age, cbrt_age, log_age, young_age, old_age, equiGen_nozero)%>% 
  write_tsv("output/201020_PBSIM/phenotypes/201020_PBSIM.generation_proxy.txt", col_names = FALSE)  
  
```

##Subsetting Purebred Angus
```{r}
sim_fam %>%
  left_join(filter(simmental, asa_nbr != 2326708)) %>%
  mutate(ASA_nbr = asa_nbr,
         dummyid = international_id) %>% 
  left_join(read_csv("data/200907_SIM/breeds.csv") %>% 
  pivot_wider(names_from = brd_cd, values_from = percent)) %>% 
  filter(AN == 1) %>% 
  select(dummyid, international_id) %>% 
  write_tsv("output/201020_ANGUS/201020_ANGUS.keepids.txt", col_names = FALSE)

sim_fam %>%
  left_join(filter(simmental, asa_nbr != 2326708)) %>%
  mutate(ASA_nbr = asa_nbr,
         dummyid = international_id) %>% 
  left_join(read_csv("data/200907_SIM/breeds.csv") %>% 
  pivot_wider(names_from = brd_cd, values_from = percent)) %>% 
  filter(AN == 1) %>%
    mutate(sqrt_age = age^0.5,
           cbrt_age = age^0.333,
           log_age = log(age),
           bc_age = bcPower(age, lambda = 0.0345),
           young_age = case_when(birth_year >= 2012 ~ age,
                                 TRUE ~ NA_real_),
           young_log_age = case_when(birth_year >= 2012 ~ log_age,
                                 TRUE ~ NA_real_),
           equiGen_nozero = case_when(equiGen <= 1 ~ NA_real_,
                                 TRUE ~ equiGen),
         old_age = case_when(birth_year < 2012 ~ age,
                               TRUE ~ NA_real_),
         old_log_age = case_when(birth_year < 2012 ~ log_age,
                               TRUE ~ NA_real_)) %>% 
  select(dummyid, international_id, age, sqrt_age, cbrt_age, log_age)%>% 
  write_tsv("output/201020_ANGUS/phenotypes/201020_ANGUS.generation_proxy.txt", col_names = FALSE)  


read_csv("data/200907_SIM/breeds.csv") %>% 
  pivot_wider(names_from = brd_cd, values_from = percent) %>% 
  filter(AN == 1) %>% 
  gg
```

### Age Summary Stats
```{r}
simmental %>% 
  select(age) %>% 
  summarize(mean_age = mean(age, na.rm = TRUE),
            median_age = median(age, na.rm = TRUE),
            sd_age = sd(age, na.rm = TRUE),
            min_age = min(age, na.rm = TRUE),
            max_age = max(age, na.rm = TRUE))
```

## Transformations to Age: {.tabset}

### Untransformed
```{r}
simmental %>% 
  select(age) %>%
  ggplot()+
  geom_histogram(aes(x = age), mutate(simmental, z = FALSE), bins = 100, fill = "dodgerblue")+
  geom_histogram(aes(x = age), mutate(simmental, z = TRUE), bins = 250, fill = "dodgerblue")+
  theme_cowplot()+
  labs(title = "Simmental Ages\n(Untransformed)", x = "Age (as of 9-15-2020)", y = "Count")+
  facet_zoom(xlim = c(10,45), ylim = c(0,175), zoom.data = z, horizontal = FALSE)
```

### Square Root
```{r}
simmental%>% 
  ggplot()+
  geom_histogram(aes(x = sqrt_age), mutate(simmental, z = FALSE), bins = 100, fill = "dodgerblue")+
  geom_histogram(aes(x = sqrt_age), mutate(simmental, z = TRUE), bins = 250, fill = "dodgerblue")+
  theme_cowplot()+
  labs(title = "Simmental Ages\n(Square Root Transformed)", x = "Age (as of 9-15-2020)", y = "Count")+
  facet_zoom(xlim = c(3,7), ylim = c(0,175), zoom.data = z, horizontal = FALSE)
```

### Cube Root
```{r}
simmental%>% 
  ggplot()+
  geom_histogram(aes(x = cbrt_age), mutate(simmental, z = FALSE), bins = 100, fill = "dodgerblue")+
  geom_histogram(aes(x = cbrt_age), mutate(simmental, z = TRUE), bins = 250, fill = "dodgerblue")+
  theme_cowplot()+
  labs(title = "Simmental Ages\n(Cube Root Transformed)", x = "Age (as of 9-15-2020)", y = "Count")+
  facet_zoom(xlim = c(2,3.6), ylim = c(0,175), zoom.data = z, horizontal = FALSE)
```

### Log 
```{r}
simmental%>% 
  ggplot()+
  geom_histogram(aes(x = log_age), mutate(simmental, z = FALSE), bins = 100, fill = "dodgerblue")+
  geom_histogram(aes(x = log_age), mutate(simmental, z = TRUE), bins = 250, fill = "dodgerblue")+
  theme_cowplot()+
  labs(title = "Simmental Ages\n(log Transformed)", x = "Age (as of 9-15-2020)", y = "Count")+
  facet_zoom(xlim = c(2,3.6), ylim = c(0,175), zoom.data = z, horizontal = FALSE)
```

### Box-Cox
```{r}
simmental%>% 
  ggplot()+
  geom_histogram(aes(x = bc_age), mutate(simmental, z = FALSE), bins = 100, fill = "dodgerblue")+
  geom_histogram(aes(x = bc_age), mutate(simmental, z = TRUE), bins = 250, fill = "dodgerblue")+
  theme_cowplot()+
  labs(title = "Simmental Ages\n(Box-Cox Transformed)", x = "Age (as of 9-15-2020)", y = "Count")+
  facet_zoom(xlim = c(2,3.6), ylim = c(0,175), zoom.data = z, horizontal = FALSE)
```


## Generation Counts

### Summary Stats for Generation Number
Unable to actually calculate this at this point as we haven't received the updated pedigree from Red Angus
```{r}
simmental %>%
  select(equiGen, fullGen, maxGen) %>%
  summarize_all(list(mean, median, sd, min, max))%>% 
  gather(key = "key", value = "value") %>%
  separate(key, c("variable", "stat"), sep = "_") %>%
  spread(stat, value) %>% 
  rename(generation_count = variable, mean = fn1, median = fn2, sd = fn3, min = fn4, max = fn5)
```

## Generation vs. Age (Simmental)

### Correlations between generation proxies:
```{r}
simmental %>% 
  select(age, birth_year, equiGen, fullGen, maxGen) %>% 
  cor()

simmental %>% 
  filter(equiGen > 1) %>% 
  select(age, birth_year, equiGen, fullGen, maxGen) %>% 
  cor()
```


```{r}
simmental %>% 
  ggplot(aes(birth_date, equiGen))+
  geom_point()+
  geom_smooth()

simmental %>% 
  ggplot(aes(birth_date, maxGen))+
  geom_point()+
  geom_smooth()

lm(maxGen~age, data = simmental) %>% summary()
```


## Distributions of Generation {.tabset}

### equiGen
Note the high number of "zero" generations here. Wondering if we shouldn't count those as NA if we choose to go this route?
```{r}
simmental%>% 
  ggplot()+
  geom_histogram(aes(x = equiGen), bins = 100, fill = "dodgerblue")+
  theme_cowplot()+
  labs(title = "Simmental Equivalent Generations", x = "Equivalent Generation", y = "Count")

```

### equiGen squared
```{r}
simmental%>% 
  ggplot()+
  geom_histogram(aes(x = equiGen^2), bins = 100, fill = "dodgerblue")+
  theme_cowplot()+
  labs(title = "Simmental Equivalent Generations Squared", x = "Equivalent Generation Squared", y = "Count")
```

### fullGen
```{r}
simmental%>% 
  ggplot()+
  geom_histogram(aes(x = fullGen),  bins = 10, fill = "dodgerblue")+
  theme_cowplot()+
  labs(title = "Simmental Full Generations", x = "Full Generation", y = "Count")
```

### maxGen
```{r}
simmental%>% 
  ggplot()+
  geom_histogram(aes(x = maxGen),  bins = 25, fill = "dodgerblue")+
  theme_cowplot()+
  labs(title = "Simmental Max Generations", x = "Maximum Generation", y = "Count")
```

# Red Angus

## Generating phenotype files

```{r}
ran_animals = 
  read_csv("data/200910_RAN/redangus/raaa.animal.20200827.csv") %>% 
    left_join(read_csv("data/200910_RAN/redangus/genomic_id.crossreference.csv")) %>% 
    left_join(read_csv("data/200910_RAN/redangus/breeder_owner_information.csv"))



redangus = 
  read_csv("output/200910_RAN/phenotypes/200910_RAN.info.csv")


#outputs age phenotype file
redangus %>% 
  select(international_id, age, birth_year) %>% 
  distinct() %>% 
  mutate(sqrt_age = age^0.5,
         cbrt_age = age^0.333,
         log_age = log(age),
         bc_age = bcPower(age, lambda = -0.237),
         young_age = case_when(birth_year >= 2012 ~ age,
                               TRUE ~ NA_real_),
         young_log_age = case_when(birth_year >= 2012 ~ log_age,
                               TRUE ~ NA_real_),
         old_age = case_when(birth_year < 2012 ~ age,
                               TRUE ~ NA_real_),
         old_log_age = case_when(birth_year < 2012 ~ log_age,
                               TRUE ~ NA_real_)) %>% 
  mutate(dummyfam = international_id) %>% 
  select(dummyfam, international_id, age, sqrt_age, cbrt_age, log_age, bc_age, young_age, young_log_age, old_age, old_log_age) %>% 
  write_tsv("output/200910_RAN/phenotypes/200910_RAN.generation_proxy.txt",
            col_names = FALSE)

read_tsv("output/200910_RAN/phenotypes/200910_RAN.generation_proxy.txt",
         col_names = c("dummyfam", "international_id", "age", "sqrt_age", "cbrt_age", "log_age", "bc_age")) %>% 
  filter(!is.na(age)) %>% 
  select(dummyfam, international_id) %>% 
  write_tsv("output/200910_RAN/200910_RAN.phenotyped_ids.txt",
            col_names = FALSE)

```


## Imputation Statistics

This is the number of variants that we're dealing with when we potentially filter on R^2 and/or MAF

R2	NO_MAF	     MAF_0.001	 MAF_0.005   MAF_0.01
________________________________
0.0	 42,156,321	 17,850,177  13,912,529	 12,758,503
0.2	 31,408,474	 15,860,556	 12,937,387  12,018,389
0.3	 29,650,214	 15,495,110	 12,768,056  11,888,636
0.4	 27,856,427	 15,143,422	 12,604,835  11,759,568
0.5	 25,877,004	 14,762,508	 12,422,846  11,611,406


the R^2 > 0.4 filter with 0.005 MAF filter would leave 12,604,835 variants


Initially my thought is that we should use 


## Exploring Generation Proxy Phenotypes 

### Calculating Generation Number
```{r, eval = FALSE}
countGen(
  data.frame(
    id = 1:5,
    dam = c(0,0,1,1,4),
    sire = c(0,0,2,2,3)
  )
)

ran_ped = 
  read_csv("data/200910_RAN/All_Animals_SireDam.csv") %>% 
  select(id = anm_key, sire = sire_key, dam = dam_key) %>% 
  replace_na(list(sire = 0, dam = 0))

ran_ped = 
  orderPed(ran_ped)

```



### Age Summary Stats
```{r}
redangus %>% 
  select(age) %>% 
  summarize(mean_age = mean(age, na.rm = TRUE),
            median_age = median(age, na.rm = TRUE),
            sd_age = sd(age, na.rm = TRUE),
            min_age = min(age, na.rm = TRUE),
            max_age = max(age, na.rm = TRUE))
```

## Transformations to Age: {.tabset}

### Untransformed
```{r}
ages = 
  redangus %>% 
    select(age) %>%
    mutate(sqrt_age = age^0.5,
           cbrt_age = age^0.333,
           log_age = log(age),
           bc_age = bcPower(age, lambda = -0.237)
           )


  
redangus %>% 
  select(age) %>%
  ggplot()+
  geom_histogram(aes(x = age), mutate(redangus, z = FALSE), bins = 100, fill = "indianred")+
  geom_histogram(aes(x = age), mutate(redangus, z = TRUE), bins = 250, fill = "indianred")+
  theme_cowplot()+
  labs(title = "Red Angus Ages\n(Untransformed)", x = "Age (as of 9-15-2020)", y = "Count")+
  facet_zoom(xlim = c(10,45), ylim = c(0,175), zoom.data = z, horizontal = FALSE)
```

### Square Root
```{r}
ages%>% 
  ggplot()+
  geom_histogram(aes(x = sqrt_age), mutate(ages, z = FALSE), bins = 100, fill = "indianred")+
  geom_histogram(aes(x = sqrt_age), mutate(ages, z = TRUE), bins = 250, fill = "indianred")+
  theme_cowplot()+
  labs(title = "Red Angus Ages\n(Square Root Transformed)", x = "Age (as of 9-15-2020)", y = "Count")+
  facet_zoom(xlim = c(3,7), ylim = c(0,175), zoom.data = z, horizontal = FALSE)
```

### Cube Root
```{r}
ages%>% 
  ggplot()+
  geom_histogram(aes(x = cbrt_age), mutate(ages, z = FALSE), bins = 100, fill = "indianred")+
  geom_histogram(aes(x = cbrt_age), mutate(ages, z = TRUE), bins = 250, fill = "indianred")+
  theme_cowplot()+
  labs(title = "Red Angus Ages\n(Cube Root Transformed)", x = "Age (as of 9-15-2020)", y = "Count")+
  facet_zoom(xlim = c(2,3.6), ylim = c(0,175), zoom.data = z, horizontal = FALSE)
```

### Log 
```{r}
ages%>% 
  ggplot()+
  geom_histogram(aes(x = log_age), mutate(ages, z = FALSE), bins = 100, fill = "indianred")+
  geom_histogram(aes(x = log_age), mutate(ages, z = TRUE), bins = 250, fill = "indianred")+
  theme_cowplot()+
  labs(title = "Red Angus Ages\n(log Transformed)", x = "Age (as of 9-15-2020)", y = "Count")+
  facet_zoom(xlim = c(2,3.6), ylim = c(0,175), zoom.data = z, horizontal = FALSE)
```

### Box-Cox
```{r}
ages%>% 
  ggplot()+
  geom_histogram(aes(x = bc_age), mutate(ages, z = FALSE), bins = 100, fill = "indianred")+
  geom_histogram(aes(x = bc_age), mutate(ages, z = TRUE), bins = 250, fill = "indianred")+
  theme_cowplot()+
  labs(title = "Red Angus Ages\n(Box-Cox Transformed)", x = "Age (as of 9-15-2020)", y = "Count")+
  facet_zoom(xlim = c(2,3.6), ylim = c(0,175), zoom.data = z, horizontal = FALSE)
```


## Summary Stats for Generation Number
Unable to actually calculate this at this point as we haven't received the updated pedigree from Red Angus
```{r}
# redangus %>% 
#   select(age) %>% 
#   summarize(mean_age = mean(age, na.rm = TRUE),
#             median_age = median(age, na.rm = TRUE),
#             sd_age = sd(age, na.rm = TRUE),
#             min_age = min(age, na.rm = TRUE),
#             max_age = max(age, na.rm = TRUE))
```



